### 2024-11-12 submission/manuscript.Rmd
.Rmd run successfully to create .tex .pdf

---
title: "**NAME OF THIS STUDY**"
bibliography: references.bib
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: header.tex
csl: mbio.csl
fontsize: 11pt
geometry: margin=1.0in
---

## ff
csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles

```{r knitr_settings, eval=TRUE, echo=FALSE, cache=FALSE}
knitr:: opts_chunk$set("tidy" = FALSE)
knitr:: opts_chunk$set("echo" = FALSE)
knitr:: opts_chunk$set("eval" = TRUE)
knitr:: opts_chunk$set("warning" = FALSE)
knitr:: opts_chunk$set("cache" = FALSE)

inline_hook <- function(x){
	print(x)

	if(is.list(x)){
		x <- unlist(x)
	}

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=1, nsmall=1, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
```

\vspace{35mm}

Running title: INSERT RUNNING TITLE HERE

\vspace{35mm}

Munkh^1, Joeseph P. Schmo^2, Sally J. Rivers^1, Patrick D. Schloss^1$\dagger$^

\vspace{40mm}

$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu

1\. Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI 48109

2\. Other department contact information

\newpage
\linenumbers

## Abstract

\newpage

## Introduction

## Results and Discussion

## Conclusions

## Materials and Methods

\newpage

Insert figure legends with the first sentence in bold, for example:

**Figure 1. Number of OTUs sampled among bacterial and archaeal 16S rRNA gene sequences for different OTU definitions and level of sequencing effort.** Rarefaction curves for different OTU definitions of Bacteria (A) and Archaea (B). Rarefaction curves for the coarse environments in Table 1 for Bacteria (C) and Archaea (D).

\newpage

## References

### 2024-11-12 submission/manuscript.tex
.Rmd run successfully to create .tex .pdf

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  11pt,
]{article}
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={NAME OF THIS STUDY},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{NAME OF THIS STUDY}}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\subsection{ff}\label{ff}

csl: mbio.csl \#Get themes at
\url{https://github.com/citation-style-language/styles}

\vspace{35mm}

Running title: INSERT RUNNING TITLE HERE

\vspace{35mm}

Munkh\^{}1, Joeseph P. Schmo\^{}2, Sally J. Rivers\^{}1, Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

\vspace{40mm}

\(\dagger\) To whom correspondence should be addressed:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

1. Department of Microbiology and Immunology, University of Michigan,
Ann Arbor, MI 48109

2. Other department contact information

\newpage
\linenumbers

\subsection{Abstract}\label{abstract}

\newpage

\subsection{Introduction}\label{introduction}

\subsection{Results and Discussion}\label{results-and-discussion}

\subsection{Conclusions}\label{conclusions}

\subsection{Materials and Methods}\label{materials-and-methods}

\newpage

Insert figure legends with the first sentence in bold, for example:

\textbf{Figure 1. Number of OTUs sampled among bacterial and archaeal
16S rRNA gene sequences for different OTU definitions and level of
sequencing effort.} Rarefaction curves for different OTU definitions of
Bacteria (A) and Archaea (B). Rarefaction curves for the coarse
environments in Table 1 for Bacteria (C) and Archaea (D).

\newpage

\subsection{References}\label{references}

\end{document}

### 2024-11-13 submission/manuscript.Rmd
draft of outline

@@
 ---
-title: "**NAME OF THIS STUDY**"
+title: "**Spatio-temporal distinct patterns in variations of $PM_{10}$ and $PM_{2.5}$ relative to the recent drivings of emission sources in Mongolia**"
 bibliography: references.bib
@@
       in_header: header.tex
-csl: mbio.csl
+csl: nature-communications.csl #Get themes at <https://github.com/citation-style-language/styles>
 fontsize: 11pt
@@
 ---
-
-## ff
-csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles
 
@@
 
-Munkh^1, Joeseph P. Schmo^2, Sally J. Rivers^1, Patrick D. Schloss^1$\dagger$^
+Munkh\^1, Joeseph P. Schmo\^2, Sally J. Rivers\^1, Patrick D. Schloss^1^$\dagger$
 
@@
 
-$\dagger$ To whom correspondence should be addressed: pschloss@umich.edu
+$\dagger$ To whom correspondence should be addressed: [pschloss\@umich.edu](mailto:pschloss@umich.edu){.email}
 
@@
 
-\newpage
+Storyline:
+
+1.  A new pattern is emerged
+2.
+Air quality in urban sites is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
+[Air quality is governed by natural dust emission, and anthropogenic emissions]
+3.  With recent growing interest in urban life style, and combustion of coal/oyutolgoi for heating winter conditions results a highly increase in not only capital city but also towns
+4.  In a result, spring coarse dust, plus winter fine pollutants
+5.
+spring coarse dust is immediately transported and deposited in the source area, whereas winter fine pollutants is permanently stayed in the source area due to stagnant atmosphere govern over entire country., perhaps floating in the near surface, deposits in the surface]
+6.
+Alarms, the Mongolian dust in the spring, optical properties might be shifted; this gives ...
+Gobi dust and sand storms has become tuiren, from the shoroon shuurga.
+which clearly requires the attention.
+7.  r ratio shows ... emission source; dust might carry anthropogenic fine particulates as well. \newpage
 
@@
 
+Significance of PMs in the climate system
+
+* PM consists of dust and aerosol, and emissions. 
+* particulate matters has a significant effects on the climate system, altering the solar incidence, cloud formation, and precipitation. 
+* concentrations of particulate matter is ephederemal, yet vary depending on whether the pollution cause is natural or industrial, local or transported, seasonal or non-seasonal, makes complex and challenging.
+
+Moreover, emerging shifts in the PM emissions and their patterns alters continuously due to their dependencies on various drivings.
+
+* It is well-informed that concentrations of air particulate matter solely depend on urbanization and economic situations to the area of the interest of the country.
+Globally, 7.3 billion people are directly exposed to unsafe average annual PM2.5 concentrations, and 80% of them living in low- and middle-income countries, where economies often rely heavily on polluting industries.
+A similar pattern of the significant disparities in air quality among income and racial/ethnic groups, as well as between urban and rural areas was reported in USA (Liu et al., 2021).
+Despite this disparity, meteorological effects such as dust storm, stagnant weather plays important role in the spatiotemporal variability of PM10 and PM2.5.
+
+Similarly, Mongolia might turn into the not only coarse-dust source region but also fine-mixed coarse-dust region.
+
+* Use of coal is increased 
+* Urbanization 
+* The monthly mean concentrations of PM10 (PM2.5) reached annual maximum in December and January due to winter synoptic governing conditions in Ulaanbaatar, capital city of Mongolia (Jugder). 
+* Despite this, the spring dust storms creates another polluted season in UB.
+On spring, the dust storm from the Gobi Desert contribute significantly to increased aerosols in the atmosphere and ambient air pollution, leading to sporadic peaks in PM10 concentrations reaching as high as 64-234 $\mu g m^{-3}$ per day or exceeding 6000 $\mu g m^{-3}$ per hour (Jugder).
+* Siberian anticyclonic activity governed over Mongolia, which create a significant vulnerability to winter air pollution in the populated areas. 
+* A such changes in PM10 and PM2.5 to stagnant weather conditions, and local or transported dust was also observed in other countries China (Wang), Korea (Kim) and Japan ().
+
+Following problems
+
+* On downwind regions
+* On national-level
+Demonstrating temporal and spatial variations of air particulate matter has become important for understanding characteristics of particulate matter in the climate system, providing valuable information for well-established air quality measures, and illustrating the good trace data for health studies.
+Because particulate pollutants have a great impact on human health (Dockery and Pope,1994; Harrison and Yin, 2000; Hong et al., 2002), high atmospheric concentrations of these pollutants was a major concern particularly in urban areas, in the last 2-3 decades.
+Recent studies highlight that even low concentrations of these pollutants can lead to various health issues, and may associate with morbidity and mortality across the life span (Zigler et al., 2017).
+Children exposed to high levels of air pollution show increased rates of asthma, decreased lung function growth, and increased risk of early markers of cardiovascular disease (Bourdrel et al., 2017; Gauderman et al., 2015; Hehua et al., 2017).
+Short-term exposure with high level of PM10 resulted the chronic cardiovascular disease in Mongolia (Enkhjargal 2020).
+In addition to these health issues, (prenatal) neurodevelopmental impacts such as effects on intelligence, attention, autism, and mood, while aging populations experience accelerated cognitive decline when exposed to high levels of pollution is detected (Power et al., 2016).
+Long-term exposure to low levels of particulate matter, such as concentrations as low as 10 $\mu g m^{-3}$ (equilibrium to WHO Air Quality Guidelines), has been linked to increased lung cancer in the EU (Hvidtfeldt et al.
+2021), with similar evidences reported in Canada (Bai et al., 2019), and significantly higher rates captured in China with concentrations up to 30 $\mu g m^{-3}$.
+Apparently, pollutants of particulate matters has effects to various health issues with the different thresholds and exposure durations.
+However, more in-depth and diversified research on air pollution and its health effects is essential, with the detailed information is necessary (Tan et al 2021) to have accuracy of assessing exposure to air pollution during developmentally relevant time periods, such as trimesters or months (Becerra et al., 2013; Gong et al., 2014; Kalkbrenner et al., 2014) or weeks (Chiu et al., 2016).
+Many research findings/Numerous research findings have advanced the field, and air quality indices is widely used for providing guidance, and public perception of air quality has been improved (Mirabelli et al., 2020).
+
+Therefore, we aimed to demonstrate the distinct temporal and spatial variations of PM2.5 and PM10 across urban and rural Mongolia using extensive data from 2008 to 2020.
+
+* The present study will contribute significantly to the understanding of air particulate matter patterns in Mongolia and providing comprehensive data insights for policymakers and public health sectors.
+* Our findings is useful not only for addressing national health impacts but also beneficial for understanding air particulate matter as ambient air pollution, and tackling atmospheric aerosol effects in the climate system, and revealing their transboundary effects to the downwind regions in South-east Asia.
+
+\newpage
+
 ## Results and Discussion
+
+[@Schloss2009] dksl ss [@Munkh2017] **Figure 3. Geographic locations of study sites** is shown in the wind speed map and elevation maps.
+
+\newpage
 
@@
 
+\newpage
+
 ## Materials and Methods
 
-\newpage
+### A description of study sites
 
+According to the spatial magnitude of wind stress in Mongolia (Figure 1), the largest magnitude of wind speed is on the Gobi sites, particularly those located in the southeast edge of the country.
+
+-   The impact of high winds on plant diversity varies across environmental gradients of precipitation and soil fertility (Milchunas et al., 1988).
+-   In the desert steppe zone, species richness was lower in the drier years but did not vary with grazing pressure.
+-   In the steppe zone, species richness varied significantly with grazing pressure but did not vary between years.
+Species richness is not impacted by grazing gradient in desert steppe, but it is in the steppe (Cheng et al., 2011).
+
+In the last 2 decades, due to poverty and natural disasters there is population immigration has taken place from the rural to urban, especially to capital city of Mongolia.
+Due to tiny infrastructure to provide the mega city with the dense population, it introduces the urban pollution.
+Therefore, Ulaanbaatar air particulate matter mainly reflects the coal burning, and partly, natural dust.
+
+Consequently, the atmospheric environment and climate for Mongolian Gobi has been impacted the most by frequent dust and and sand storm in the spring.
+
+Our study was carried out in Dalanzadgad (town center) (Tbl.
+1; 43.57°N, 104.42°E), Sainshand (Tbl.
+1; 44.87°N, 110.12°E) and Zamyn-Uud (Tbl.
+1; 43.72°N, 111.90°E) in the Gobi Desert, and at Ulaanbaatar (Tbl.??.??°N, 104.42°E) (city center) located in the temperate Mongolian steppe of Mongolia (Figure 2).
+Nomads and settlements of this sum have raised a large number of livestock, and they rank at number 30 out of 329 sums for the largest number of livestock raised per sum (Saizen et al., 2010).
+In the last decade, the number of dust events associated with wind erodibility increased by 30 % in Bayan-Önjüül (Kurosaki et al., 2011).
+This is an area where dust emissions activity has been monitored on a long-term basis (Shinoda et al., 2010a) at a dust observation site (DOS) adjacent to the study site (Fig.
+1a).
+According to long-term meteorological observations made at the monitoring station of the Institute of Meteorology and Hydrology of Mongolia located near the site, the prevailing wind direction is northwest.
+Mean annual precipitation is 163 mm, and mean temperature is 0.1◦C for the period 1995 to 2005 (Shinoda et al., 2010b).
+Soil texture is dominated by sand (98.1 %, with only 1.3 % clay and 0.6 % silt; Table 1; Shinoda et al., 2010a).
 Insert figure legends with the first sentence in bold, for example:
 
-**Figure 1.
-Number of OTUs sampled among bacterial and archaeal 16S rRNA gene sequences for different OTU definitions and level of sequencing effort.** Rarefaction curves for different OTU definitions of Bacteria (A) and Archaea (B).
-Rarefaction curves for the coarse environments in Table 1 for Bacteria (C) and Archaea (D).
+**Figure 1. Geographic locations of study sites** is shown in the wind speed map and elevation maps.
 
-\newpage
+**Table 1. A description of datasets obtained at the sites**
+
+**Figure 2.** is shown in the wind speed map and elevation maps.
+
+\cite{Schloss2009} \newpage
 

### 2024-11-15 submission/manuscript.Rmd
outline

@@
 ---
-title: "**Spatio-temporal distinct patterns in variations of $PM_{10}$ and $PM_{2.5}$ relative to the recent drivings of emission sources in Mongolia**"
+title: "**Spatio-temporal distinct patterns in variations of $PM_{10}$ and $PM_{2.5}$
+  relative to the recent drivings of emission sources in Mongolia**"
 bibliography: references.bib
@@
       in_header: header.tex
-csl: nature-communications.csl #Get themes at <https://github.com/citation-style-language/styles>
+  html_document:
+    df_print: paged
+csl: "nature-communications.csl"
 fontsize: 11pt
 geometry: margin=1.0in
+editor_options:
+  markdown:
+    wrap: 72
 ---
@@
 
-Munkh\^1, Joeseph P. Schmo\^2, Sally J. Rivers\^1, Patrick D. Schloss^1^$\dagger$
+Munkh\^1, Joeseph P. Schmo\^2, Sally J. Rivers\^1, Patrick D.
+Schloss^1^$\dagger$
 
@@
 
-$\dagger$ To whom correspondence should be addressed: [pschloss\@umich.edu](mailto:pschloss@umich.edu){.email}
+$\dagger$ To whom correspondence should be addressed:
+[pschloss\@umich.edu](mailto:pschloss@umich.edu){.email}
 
-1\. Department of Microbiology and Immunology, University of Michigan, Ann Arbor, MI 48109
+1\. Department of Microbiology and Immunology, University of Michigan,
+Ann Arbor, MI 48109
 
@@
 1.  A new pattern is emerged
-2.
-Air quality in urban sites is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
-[Air quality is governed by natural dust emission, and anthropogenic emissions]
-3.  With recent growing interest in urban life style, and combustion of coal/oyutolgoi for heating winter conditions results a highly increase in not only capital city but also towns
+2.  Air quality in urban sites is episodically dictated by dust events
+    in spring or late autumn, yet seasonally governed by anthropogenic
+    emissions in winter. [Air quality is governed by natural dust
+    emission, and anthropogenic emissions]
+3.  With recent growing interest in urban life style, and combustion of
+    coal/oyutolgoi for heating winter conditions results a highly
+    increase in not only capital city but also towns
 4.  In a result, spring coarse dust, plus winter fine pollutants
-5.
-spring coarse dust is immediately transported and deposited in the source area, whereas winter fine pollutants is permanently stayed in the source area due to stagnant atmosphere govern over entire country., perhaps floating in the near surface, deposits in the surface]
-6.
-Alarms, the Mongolian dust in the spring, optical properties might be shifted; this gives ...
-Gobi dust and sand storms has become tuiren, from the shoroon shuurga.
-which clearly requires the attention.
-7.  r ratio shows ... emission source; dust might carry anthropogenic fine particulates as well. \newpage
+5.  spring coarse dust is immediately transported and deposited in the
+    source area, whereas winter fine pollutants is permanently stayed in
+    the source area due to stagnant atmosphere govern over entire
+    country., perhaps floating in the near surface, deposits in the
+    surface]
+6.  Alarms, the Mongolian dust in the spring, optical properties might
+    be shifted; this gives ... Gobi dust and sand storms has become
+    tuiren, from the shoroon shuurga. which clearly requires the
+    attention.
+7.  r ratio shows ... emission source; dust might carry anthropogenic
+    fine particulates as well. \newpage
 
@@
 
-Significance of PMs in the climate system
+Mongolian dust is well-known by its term of coarse, \* color of brown,
+\* its influence on downwind regions, \* and big role in global dust, \*
+and its seasonality.
 
-* PM consists of dust and aerosol, and emissions. 
-* particulate matters has a significant effects on the climate system, altering the solar incidence, cloud formation, and precipitation. 
-* concentrations of particulate matter is ephederemal, yet vary depending on whether the pollution cause is natural or industrial, local or transported, seasonal or non-seasonal, makes complex and challenging.
+Because, it mainly consists of coarse fractions researchers neglect its
+role in global climate into elaborating global dust-aerosol effects.
+*However, recent change in the driving of the emissions of particulate
+matters might cause fluctuations* its spatio-temporal variations and \*
+characteristics and \* emerge a new pattern, eventually it may lead to
+the role in the regional and global dust, so on climate system through
+\* altering solar incidence, cloud formation, and precipitation or
+warming \* the climate with the fraction of finer particulates.
 
-Moreover, emerging shifts in the PM emissions and their patterns alters continuously due to their dependencies on various drivings.
+Mongolia might turn into the not only coarse-dust source region but also
+fine-mixed coarse-dust region.
 
-* It is well-informed that concentrations of air particulate matter solely depend on urbanization and economic situations to the area of the interest of the country.
-Globally, 7.3 billion people are directly exposed to unsafe average annual PM2.5 concentrations, and 80% of them living in low- and middle-income countries, where economies often rely heavily on polluting industries.
-A similar pattern of the significant disparities in air quality among income and racial/ethnic groups, as well as between urban and rural areas was reported in USA (Liu et al., 2021).
-Despite this disparity, meteorological effects such as dust storm, stagnant weather plays important role in the spatiotemporal variability of PM10 and PM2.5.
+-   Use of coal is increased
 
-Similarly, Mongolia might turn into the not only coarse-dust source region but also fine-mixed coarse-dust region.
+-   Urbanization
 
-* Use of coal is increased 
-* Urbanization 
-* The monthly mean concentrations of PM10 (PM2.5) reached annual maximum in December and January due to winter synoptic governing conditions in Ulaanbaatar, capital city of Mongolia (Jugder). 
-* Despite this, the spring dust storms creates another polluted season in UB.
-On spring, the dust storm from the Gobi Desert contribute significantly to increased aerosols in the atmosphere and ambient air pollution, leading to sporadic peaks in PM10 concentrations reaching as high as 64-234 $\mu g m^{-3}$ per day or exceeding 6000 $\mu g m^{-3}$ per hour (Jugder).
-* Siberian anticyclonic activity governed over Mongolia, which create a significant vulnerability to winter air pollution in the populated areas. 
-* A such changes in PM10 and PM2.5 to stagnant weather conditions, and local or transported dust was also observed in other countries China (Wang), Korea (Kim) and Japan ().
+-   The monthly mean concentrations of PM10 (PM2.5) reached annual
+    maximum in December and January due to winter synoptic governing
+    conditions in Ulaanbaatar, capital city of Mongolia (Jugder).
 
-Following problems
+-   Despite this, the spring dust storms creates another polluted season
+    in UB. On spring, the dust storm from the Gobi Desert contribute
+    significantly to increased aerosols in the atmosphere and ambient
+    air pollution, leading to sporadic peaks in PM10 concentrations
+    reaching as high as 64-234 $\mu g m^{-3}$ per day or exceeding 6000
+    $\mu g m^{-3}$ per hour (Jugder).
 
-* On downwind regions
-* On national-level
-Demonstrating temporal and spatial variations of air particulate matter has become important for understanding characteristics of particulate matter in the climate system, providing valuable information for well-established air quality measures, and illustrating the good trace data for health studies.
-Because particulate pollutants have a great impact on human health (Dockery and Pope,1994; Harrison and Yin, 2000; Hong et al., 2002), high atmospheric concentrations of these pollutants was a major concern particularly in urban areas, in the last 2-3 decades.
-Recent studies highlight that even low concentrations of these pollutants can lead to various health issues, and may associate with morbidity and mortality across the life span (Zigler et al., 2017).
-Children exposed to high levels of air pollution show increased rates of asthma, decreased lung function growth, and increased risk of early markers of cardiovascular disease (Bourdrel et al., 2017; Gauderman et al., 2015; Hehua et al., 2017).
-Short-term exposure with high level of PM10 resulted the chronic cardiovascular disease in Mongolia (Enkhjargal 2020).
-In addition to these health issues, (prenatal) neurodevelopmental impacts such as effects on intelligence, attention, autism, and mood, while aging populations experience accelerated cognitive decline when exposed to high levels of pollution is detected (Power et al., 2016).
-Long-term exposure to low levels of particulate matter, such as concentrations as low as 10 $\mu g m^{-3}$ (equilibrium to WHO Air Quality Guidelines), has been linked to increased lung cancer in the EU (Hvidtfeldt et al.
-2021), with similar evidences reported in Canada (Bai et al., 2019), and significantly higher rates captured in China with concentrations up to 30 $\mu g m^{-3}$.
-Apparently, pollutants of particulate matters has effects to various health issues with the different thresholds and exposure durations.
-However, more in-depth and diversified research on air pollution and its health effects is essential, with the detailed information is necessary (Tan et al 2021) to have accuracy of assessing exposure to air pollution during developmentally relevant time periods, such as trimesters or months (Becerra et al., 2013; Gong et al., 2014; Kalkbrenner et al., 2014) or weeks (Chiu et al., 2016).
-Many research findings/Numerous research findings have advanced the field, and air quality indices is widely used for providing guidance, and public perception of air quality has been improved (Mirabelli et al., 2020).
+-   Siberian anticyclonic activity governed over Mongolia, which create
+    a significant vulnerability to winter air pollution in the populated
+    areas.
 
-Therefore, we aimed to demonstrate the distinct temporal and spatial variations of PM2.5 and PM10 across urban and rural Mongolia using extensive data from 2008 to 2020.
+-   A such changes in PM10 and PM2.5 to stagnant weather conditions, and
+    local or transported dust was also observed in other countries China
+    (Wang), Korea (Kim) and Japan ().
 
-* The present study will contribute significantly to the understanding of air particulate matter patterns in Mongolia and providing comprehensive data insights for policymakers and public health sectors.
-* Our findings is useful not only for addressing national health impacts but also beneficial for understanding air particulate matter as ambient air pollution, and tackling atmospheric aerosol effects in the climate system, and revealing their transboundary effects to the downwind regions in South-east Asia.
+-   deposition
+
+-   concentrations of particulate matter is ephederemal, yet vary
+    depending on whether the pollution cause is natural or industrial,
+    local or transported, seasonal or non-seasonal, makes complex and
+    challenging.
+
+Therefore, we aimed to demonstrate the distinct temporal and spatial
+variations of PM2.5 and PM10 across urban and rural Mongolia using
+extensive data from 2008 to 2020.
+
+-   The present study will contribute significantly to the understanding
+    of air particulate matter patterns in Mongolia and providing
+    comprehensive data insights for policymakers and public health
+    sectors.
+-   Our findings is useful not only for addressing national health
+    impacts but also beneficial for understanding air particulate matter
+    as ambient air pollution, and tackling atmospheric aerosol effects
+    in the climate system, and revealing their transboundary effects to
+    the downwind regions in South-east Asia.
 
@@
 
-[@Schloss2009] dksl ss [@Munkh2017] **Figure 3. Geographic locations of study sites** is shown in the wind speed map and elevation maps.
+\newpage
+\subsection{Urban and rural distinct Spatio-temporal diverse variations of $PM_{10}$ and $PM_{2.5}$}
+
+![Distinct concentrations of coarse and fine particulates among
+sites](images/figure_3.png){width="300"}
+
+1.  Compare the concentrations of PMs at UB is the 2. Significance level
+    difference 3. Conclude
+
+\newpage
+
+![Annual variations of \$PM\_{10}\$ and
+\$PM\_{2.5}\$](images/figure_4.png){width="330"}
+
+ 1. Clear annual variations at UB and DZ from pm2.5 pollutions 2. at ZU,
+and SS has a seasonally peaks episodic spring and late autumn from PM10
+
+\newpage
+
+![Daily variations of $PM_{10}$ and $PM_{2.5}$ at UB and DZ
+sites](images/figure_5.png){width="300"}
+
+\newpage
+\subsection{Meteorological influence on $PM_{10}$ and $PM_{2.5}$ variations}
+\label{subsec2}
+
+![Relationships between meteorological major factors and variations of
+$PM_{10}$ and $PM_{2.5}$](images/figure_6.png){width="300"}
+
+\newpage
+
+![Spatio-temporal distinct feature of variations of $PM_{10}$ and
+$PM_{2.5}$ with PCA analysis](images/figure_7.png){width="200"}
+
+![Patterns of meteorology and PMs at the 4
+sites](images/figure_7b.png){width="200"}
+
+\newpage
+\subsection{Trends}
+
+![Interannual and seasonal trends of $PM_{10}$ and $PM_{2.5}$
+variations](images/figure_8.png){width="300"}
 
@@
 ## Conclusions
+
+In this study, we investigated the temporal variations of PM2.5 and PM10
+concentrations at the 4 sites of rural and urban those located along the
+the wind corridor. Three distinct variations has been detected.
+
+\begin{enumerate}
+    \item Air quality in urban sites is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
+    \item Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
+    \item Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
+\end{enumerate}
+
+A clear seasonal variations in the sites of UB and DZ is [Air quality is
+governed by natural dust emission, and anthropogenic emissions] \* Due
+to rapid increase in urban, and combustion of coal/oyutolgoi for heating
+winter conditions results a highly increase in not only capital city but
+also towns \* In a result, spring coarse dust, plus winter fine
+pollutants [spring coarse dust is immediately transported and deposited
+in the source area, whereas winter fine pollutants is permanently stayed
+in the source area due to stagnant atmosphere govern over entire
+country., perhaps float- ing in the near surface, deposits in the
+surface] \* Alarms, the Mongolian dust in the spring, optical properties
+will be shifted; this gives ... Gobi dust and sand storms has become
+tuiren, from the shoroon shuurga. which clearly requires the attention.
+
+Following problems
+
+-   On downwind regions
+-   On national-level Demonstrating temporal and spatial variations of
+    air particulate matter has become important for understanding
+    characteristics of particulate matter in the climate system,
+    providing valuable information for well-established air quality
+    measures, and illustrating the good trace data for health studies.
+    Because particulate pollutants have a great impact on human health
+    (Dockery and Pope,1994; Harrison and Yin, 2000; Hong et al., 2002),
+    high atmospheric concentrations of these pollutants was a major
+    concern particularly in urban areas, in the last 2-3 decades. Recent
+    studies highlight that even low concentrations of these pollutants
+    can lead to various health issues, and may associate with morbidity
+    and mortality across the life span (Zigler et al., 2017). Children
+    exposed to high levels of air pollution show increased rates of
+    asthma, decreased lung function growth, and increased risk of early
+    markers of cardiovascular disease (Bourdrel et al., 2017; Gauderman
+    et al., 2015; Hehua et al., 2017). Short-term exposure with high
+    level of PM10 resulted the chronic cardiovascular disease in
+    Mongolia (Enkhjargal 2020). In addition to these health issues,
+    (prenatal) neurodevelopmental impacts such as effects on
+    intelligence, attention, autism, and mood, while aging populations
+    experience accelerated cognitive decline when exposed to high levels
+    of pollution is detected (Power et al., 2016). Long-term exposure to
+    low levels of particulate matter, such as concentrations as low as
+    10 $\mu g m^{-3}$ (equilibrium to WHO Air Quality Guidelines), has
+    been linked to increased lung cancer in the EU (Hvidtfeldt et al.
+    2021), with similar evidences reported in Canada (Bai et al., 2019),
+    and significantly higher rates captured in China with concentrations
+    up to 30 $\mu g m^{-3}$. Apparently, pollutants of particulate
+    matters has effects to various health issues with the different
+    thresholds and exposure durations. However, more in-depth and
+    diversified research on air pollution and its health effects is
+    essential, with the detailed information is necessary (Tan et al
+    2021) to have accuracy of assessing exposure to air pollution during
+          developmentally relevant time periods, such as trimesters or
+          months (Becerra et al., 2013; Gong et al., 2014; Kalkbrenner
+          et al., 2014) or weeks (Chiu et al., 2016). Many research
+          findings/Numerous research findings have advanced the field,
+          and air quality indices is widely used for providing guidance,
+          and public perception of air quality has been improved
+          (Mirabelli et al., 2020).
 
@@
 
-According to the spatial magnitude of wind stress in Mongolia (Figure 1), the largest magnitude of wind speed is on the Gobi sites, particularly those located in the southeast edge of the country.
+According to the spatial magnitude of wind stress in Mongolia (Figure
+1), the largest magnitude of wind speed is on the Gobi sites,
+particularly those located in the southeast edge of the country.
 
--   The impact of high winds on plant diversity varies across environmental gradients of precipitation and soil fertility (Milchunas et al., 1988).
--   In the desert steppe zone, species richness was lower in the drier years but did not vary with grazing pressure.
--   In the steppe zone, species richness varied significantly with grazing pressure but did not vary between years.
-Species richness is not impacted by grazing gradient in desert steppe, but it is in the steppe (Cheng et al., 2011).
+-   The impact of high winds on plant diversity varies across
+    environmental gradients of precipitation and soil fertility
+    (Milchunas et al., 1988).
+-   In the desert steppe zone, species richness was lower in the drier
+    years but did not vary with grazing pressure.
+-   In the steppe zone, species richness varied significantly with
+    grazing pressure but did not vary between years. Species richness is
+    not impacted by grazing gradient in desert steppe, but it is in the
+    steppe (Cheng et al., 2011).
 
-In the last 2 decades, due to poverty and natural disasters there is population immigration has taken place from the rural to urban, especially to capital city of Mongolia.
-Due to tiny infrastructure to provide the mega city with the dense population, it introduces the urban pollution.
-Therefore, Ulaanbaatar air particulate matter mainly reflects the coal burning, and partly, natural dust.
+In the last 2 decades, due to poverty and natural disasters there is
+population immigration has taken place from the rural to urban,
+especially to capital city of Mongolia. Due to tiny infrastructure to
+provide the mega city with the dense population, it introduces the urban
+pollution. Therefore, Ulaanbaatar air particulate matter mainly reflects
+the coal burning, and partly, natural dust.
 
-Consequently, the atmospheric environment and climate for Mongolian Gobi has been impacted the most by frequent dust and and sand storm in the spring.
+Consequently, the atmospheric environment and climate for Mongolian Gobi
+has been impacted the most by frequent dust and and sand storm in the
+spring.
 
-Our study was carried out in Dalanzadgad (town center) (Tbl.
-1; 43.57°N, 104.42°E), Sainshand (Tbl.
-1; 44.87°N, 110.12°E) and Zamyn-Uud (Tbl.
-1; 43.72°N, 111.90°E) in the Gobi Desert, and at Ulaanbaatar (Tbl.??.??°N, 104.42°E) (city center) located in the temperate Mongolian steppe of Mongolia (Figure 2).
-Nomads and settlements of this sum have raised a large number of livestock, and they rank at number 30 out of 329 sums for the largest number of livestock raised per sum (Saizen et al., 2010).
-In the last decade, the number of dust events associated with wind erodibility increased by 30 % in Bayan-Önjüül (Kurosaki et al., 2011).
-This is an area where dust emissions activity has been monitored on a long-term basis (Shinoda et al., 2010a) at a dust observation site (DOS) adjacent to the study site (Fig.
-1a).
-According to long-term meteorological observations made at the monitoring station of the Institute of Meteorology and Hydrology of Mongolia located near the site, the prevailing wind direction is northwest.
-Mean annual precipitation is 163 mm, and mean temperature is 0.1◦C for the period 1995 to 2005 (Shinoda et al., 2010b).
-Soil texture is dominated by sand (98.1 %, with only 1.3 % clay and 0.6 % silt; Table 1; Shinoda et al., 2010a).
-Insert figure legends with the first sentence in bold, for example:
+Our study was carried out in Dalanzadgad (town center) (Tbl. 1; 43.57°N,
+104.42°E), Sainshand (Tbl. 1; 44.87°N, 110.12°E) and Zamyn-Uud (Tbl. 1;
+43.72°N, 111.90°E) in the Gobi Desert, and at Ulaanbaatar (Tbl.??.??°N,
+104.42°E) (city center) located in the temperate Mongolian steppe of
+Mongolia (Figure 2). Nomads and settlements of this sum have raised a
+large number of livestock, and they rank at number 30 out of 329 sums
+for the largest number of livestock raised per sum (Saizen et al.,
+2010). In the last decade, the number of dust events associated with
+wind erodibility increased by 30 % in Bayan-Önjüül (Kurosaki et al.,
+2011). This is an area where dust emissions activity has been monitored
+on a long-term basis (Shinoda et al., 2010a) at a dust observation site
+(DOS) adjacent to the study site (Fig. 1a). According to long-term
+meteorological observations made at the monitoring station of the
+Institute of Meteorology and Hydrology of Mongolia located near the
+site, the prevailing wind direction is northwest. Mean annual
+precipitation is 163 mm, and mean temperature is 0.1◦C for the period
+1995 to 2005 (Shinoda et al., 2010b). Soil texture is dominated by sand
+(98.1 %, with only 1.3 % clay and 0.6 % silt; Table 1; Shinoda et al.,
+2010a). Insert figure legends with the first sentence in bold, for
+example:
 
-**Figure 1. Geographic locations of study sites** is shown in the wind speed map and elevation maps.
+\newpage
 
-**Table 1. A description of datasets obtained at the sites**
+![Geographic locations of study
+sites](images/figure_1.png){width="300"}]
 
-**Figure 2.** is shown in the wind speed map and elevation maps.
+![**Table 1**. A description of datasets obtained at the
+sites](images/table_1.png){width="400"}
 
-\cite{Schloss2009} \newpage
+\newpage
+
+![Scheme 1. Data handling procedure](images/scheme_1.png){width="300"}
+
+\newpage
+
+![Figure 2. Data gap filling](images/figure_2b.png){width="300"}
+
+![Figure 2b. Data gap filling](images/figure_2c.png){width="300"}
+
+\newpage
 

### 2024-11-15 submission/manuscript.tex
outline

@@
 \hypersetup{
-  pdftitle={NAME OF THIS STUDY},
+  pdftitle={Spatio-temporal distinct patterns in variations of PM\_\{10\} and PM\_\{2.5\} relative to the recent drivings of emission sources in Mongolia},
   hidelinks,
@@
 
-\title{\textbf{NAME OF THIS STUDY}}
+\title{\textbf{Spatio-temporal distinct patterns in variations of
+\(PM_{10}\) and \(PM_{2.5}\) relative to the recent drivings of emission
+sources in Mongolia}}
 \author{}
@@
 \maketitle
-
-\subsection{ff}\label{ff}
-
-csl: mbio.csl \#Get themes at
-\url{https://github.com/citation-style-language/styles}
 
@@
 Munkh\^{}1, Joeseph P. Schmo\^{}2, Sally J. Rivers\^{}1, Patrick D.
-Schloss\textsuperscript{1\(\dagger\)}
+Schloss\textsuperscript{1}\(\dagger\)
 
@@
 
-\newpage
+Storyline:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  A new pattern is emerged
+\item
+  Air quality in urban sites is episodically dictated by dust events in
+  spring or late autumn, yet seasonally governed by anthropogenic
+  emissions in winter. {[}Air quality is governed by natural dust
+  emission, and anthropogenic emissions{]}
+\item
+  With recent growing interest in urban life style, and combustion of
+  coal/oyutolgoi for heating winter conditions results a highly increase
+  in not only capital city but also towns
+\item
+  In a result, spring coarse dust, plus winter fine pollutants
+\item
+  spring coarse dust is immediately transported and deposited in the
+  source area, whereas winter fine pollutants is permanently stayed in
+  the source area due to stagnant atmosphere govern over entire
+  country., perhaps floating in the near surface, deposits in the
+  surface{]}
+\item
+  Alarms, the Mongolian dust in the spring, optical properties might be
+  shifted; this gives \ldots{} Gobi dust and sand storms has become
+  tuiren, from the shoroon shuurga. which clearly requires the
+  attention.
+\item
+  r ratio shows \ldots{} emission source; dust might carry anthropogenic
+  fine particulates as well. \newpage
+\end{enumerate}
 
@@
 
+Mongolian dust is well-known by its term of coarse, * color of brown, *
+its influence on downwind regions, * and big role in global dust, * and
+its seasonality.
+
+Because, it mainly consists of coarse fractions researchers neglect its
+role in global climate into elaborating global dust-aerosol effects.
+\emph{However, recent change in the driving of the emissions of
+particulate matters might cause fluctuations} its spatio-temporal
+variations and * characteristics and * emerge a new pattern, eventually
+it may lead to the role in the regional and global dust, so on climate
+system through * altering solar incidence, cloud formation, and
+precipitation or warming * the climate with the fraction of finer
+particulates.
+
+Mongolia might turn into the not only coarse-dust source region but also
+fine-mixed coarse-dust region.
+
+\begin{itemize}
+\item
+  Use of coal is increased
+\item
+  Urbanization
+\item
+  The monthly mean concentrations of PM10 (PM2.5) reached annual maximum
+  in December and January due to winter synoptic governing conditions in
+  Ulaanbaatar, capital city of Mongolia (Jugder).
+\item
+  Despite this, the spring dust storms creates another polluted season
+  in UB. On spring, the dust storm from the Gobi Desert contribute
+  significantly to increased aerosols in the atmosphere and ambient air
+  pollution, leading to sporadic peaks in PM10 concentrations reaching
+  as high as 64-234 \(\mu g m^{-3}\) per day or exceeding 6000
+  \(\mu g m^{-3}\) per hour (Jugder).
+\item
+  Siberian anticyclonic activity governed over Mongolia, which create a
+  significant vulnerability to winter air pollution in the populated
+  areas.
+\item
+  A such changes in PM10 and PM2.5 to stagnant weather conditions, and
+  local or transported dust was also observed in other countries China
+  (Wang), Korea (Kim) and Japan ().
+\item
+  deposition
+\item
+  concentrations of particulate matter is ephederemal, yet vary
+  depending on whether the pollution cause is natural or industrial,
+  local or transported, seasonal or non-seasonal, makes complex and
+  challenging.
+\end{itemize}
+
+Therefore, we aimed to demonstrate the distinct temporal and spatial
+variations of PM2.5 and PM10 across urban and rural Mongolia using
+extensive data from 2008 to 2020.
+
+\begin{itemize}
+\tightlist
+\item
+  The present study will contribute significantly to the understanding
+  of air particulate matter patterns in Mongolia and providing
+  comprehensive data insights for policymakers and public health
+  sectors.
+\item
+  Our findings is useful not only for addressing national health impacts
+  but also beneficial for understanding air particulate matter as
+  ambient air pollution, and tackling atmospheric aerosol effects in the
+  climate system, and revealing their transboundary effects to the
+  downwind regions in South-east Asia.
+\end{itemize}
+
+\newpage
+
 \subsection{Results and Discussion}\label{results-and-discussion}
+
+\newpage
+\subsection{Urban and rural distinct Spatio-temporal diverse variations of $PM_{10}$ and $PM_{2.5}$}
+
+\begin{figure}
+\centering
+\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_3.png}
+\caption{Distinct concentrations of coarse and fine particulates among
+sites}
+\end{figure}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Compare the concentrations of PMs at UB is the 2. Significance level
+  difference 3. Conclude
+\end{enumerate}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics[width=3.4375in,height=\textheight,keepaspectratio]{images/figure_4.png}
+\caption{Annual variations of \$PM\_\{10\}\$ and \$PM\_\{2.5\}\$}
+\end{figure}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Clear annual variations at UB and DZ from pm2.5 pollutions 2. at ZU,
+  and SS has a seasonally peaks episodic spring and late autumn from
+  PM10
+\end{enumerate}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_5.png}
+\caption{Daily variations of \(PM_{10}\) and \(PM_{2.5}\) at UB and DZ
+sites}
+\end{figure}
+
+\newpage
+\subsection{Meteorological influence on $PM_{10}$ and $PM_{2.5}$ variations}
+\label{subsec2}
+
+\begin{figure}
+\centering
+\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_6.png}
+\caption{Relationships between meteorological major factors and
+variations of \(PM_{10}\) and \(PM_{2.5}\)}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics[width=2.08333in,height=\textheight,keepaspectratio]{images/figure_7.png}
+\caption{Spatio-temporal distinct feature of variations of \(PM_{10}\)
+and \(PM_{2.5}\) with PCA analysis}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=2.08333in,height=\textheight,keepaspectratio]{images/figure_7b.png}
+\caption{Patterns of meteorology and PMs at the 4 sites}
+\end{figure}
+
+\newpage
+\subsection{Trends}
+
+\begin{figure}
+\centering
+\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_8.png}
+\caption{Interannual and seasonal trends of \(PM_{10}\) and \(PM_{2.5}\)
+variations}
+\end{figure}
+
+\newpage
 
@@
 
-\subsection{Materials and Methods}\label{materials-and-methods}
+In this study, we investigated the temporal variations of PM2.5 and PM10
+concentrations at the 4 sites of rural and urban those located along the
+the wind corridor. Three distinct variations has been detected.
+
+\begin{enumerate}
+    \item Air quality in urban sites is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
+    \item Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
+    \item Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
+\end{enumerate}
+
+A clear seasonal variations in the sites of UB and DZ is {[}Air quality
+is governed by natural dust emission, and anthropogenic emissions{]} *
+Due to rapid increase in urban, and combustion of coal/oyutolgoi for
+heating winter conditions results a highly increase in not only capital
+city but also towns * In a result, spring coarse dust, plus winter fine
+pollutants {[}spring coarse dust is immediately transported and
+deposited in the source area, whereas winter fine pollutants is
+permanently stayed in the source area due to stagnant atmosphere govern
+over entire country., perhaps float- ing in the near surface, deposits
+in the surface{]} * Alarms, the Mongolian dust in the spring, optical
+properties will be shifted; this gives \ldots{} Gobi dust and sand
+storms has become tuiren, from the shoroon shuurga. which clearly
+requires the attention.
+
+Following problems
+
+\begin{itemize}
+\tightlist
+\item
+  On downwind regions
+\item
+  On national-level Demonstrating temporal and spatial variations of air
+  particulate matter has become important for understanding
+  characteristics of particulate matter in the climate system, providing
+  valuable information for well-established air quality measures, and
+  illustrating the good trace data for health studies. Because
+  particulate pollutants have a great impact on human health (Dockery
+  and Pope,1994; Harrison and Yin, 2000; Hong et al., 2002), high
+  atmospheric concentrations of these pollutants was a major concern
+  particularly in urban areas, in the last 2-3 decades. Recent studies
+  highlight that even low concentrations of these pollutants can lead to
+  various health issues, and may associate with morbidity and mortality
+  across the life span (Zigler et al., 2017). Children exposed to high
+  levels of air pollution show increased rates of asthma, decreased lung
+  function growth, and increased risk of early markers of cardiovascular
+  disease (Bourdrel et al., 2017; Gauderman et al., 2015; Hehua et al.,
+  2017). Short-term exposure with high level of PM10 resulted the
+  chronic cardiovascular disease in Mongolia (Enkhjargal 2020). In
+  addition to these health issues, (prenatal) neurodevelopmental impacts
+  such as effects on intelligence, attention, autism, and mood, while
+  aging populations experience accelerated cognitive decline when
+  exposed to high levels of pollution is detected (Power et al., 2016).
+  Long-term exposure to low levels of particulate matter, such as
+  concentrations as low as 10 \(\mu g m^{-3}\) (equilibrium to WHO Air
+  Quality Guidelines), has been linked to increased lung cancer in the
+  EU (Hvidtfeldt et al. 2021), with similar evidences reported in Canada
+  (Bai et al., 2019), and significantly higher rates captured in China
+  with concentrations up to 30 \(\mu g m^{-3}\). Apparently, pollutants
+  of particulate matters has effects to various health issues with the
+  different thresholds and exposure durations. However, more in-depth
+  and diversified research on air pollution and its health effects is
+  essential, with the detailed information is necessary (Tan et al
+
+  \begin{enumerate}
+  \def\labelenumi{\arabic{enumi})}
+  \setcounter{enumi}{2020}
+  \tightlist
+  \item
+    to have accuracy of assessing exposure to air pollution during
+    developmentally relevant time periods, such as trimesters or months
+    (Becerra et al., 2013; Gong et al., 2014; Kalkbrenner et al., 2014)
+    or weeks (Chiu et al., 2016). Many research findings/Numerous
+    research findings have advanced the field, and air quality indices
+    is widely used for providing guidance, and public perception of air
+    quality has been improved (Mirabelli et al., 2020).
+  \end{enumerate}
+\end{itemize}
 
@@
 
-Insert figure legends with the first sentence in bold, for example:
+\subsection{Materials and Methods}\label{materials-and-methods}
 
-\textbf{Figure 1. Number of OTUs sampled among bacterial and archaeal
-16S rRNA gene sequences for different OTU definitions and level of
-sequencing effort.} Rarefaction curves for different OTU definitions of
-Bacteria (A) and Archaea (B). Rarefaction curves for the coarse
-environments in Table 1 for Bacteria (C) and Archaea (D).
+\subsubsection{A description of study
+sites}\label{a-description-of-study-sites}
+
+According to the spatial magnitude of wind stress in Mongolia (Figure
+1), the largest magnitude of wind speed is on the Gobi sites,
+particularly those located in the southeast edge of the country.
+
+\begin{itemize}
+\tightlist
+\item
+  The impact of high winds on plant diversity varies across
+  environmental gradients of precipitation and soil fertility (Milchunas
+  et al., 1988).
+\item
+  In the desert steppe zone, species richness was lower in the drier
+  years but did not vary with grazing pressure.
+\item
+  In the steppe zone, species richness varied significantly with grazing
+  pressure but did not vary between years. Species richness is not
+  impacted by grazing gradient in desert steppe, but it is in the steppe
+  (Cheng et al., 2011).
+\end{itemize}
+
+In the last 2 decades, due to poverty and natural disasters there is
+population immigration has taken place from the rural to urban,
+especially to capital city of Mongolia. Due to tiny infrastructure to
+provide the mega city with the dense population, it introduces the urban
+pollution. Therefore, Ulaanbaatar air particulate matter mainly reflects
+the coal burning, and partly, natural dust.
+
+Consequently, the atmospheric environment and climate for Mongolian Gobi
+has been impacted the most by frequent dust and and sand storm in the
+spring.
+
+Our study was carried out in Dalanzadgad (town center) (Tbl. 1; 43.57°N,
+104.42°E), Sainshand (Tbl. 1; 44.87°N, 110.12°E) and Zamyn-Uud (Tbl. 1;
+43.72°N, 111.90°E) in the Gobi Desert, and at Ulaanbaatar (Tbl.??.??°N,
+104.42°E) (city center) located in the temperate Mongolian steppe of
+Mongolia (Figure 2). Nomads and settlements of this sum have raised a
+large number of livestock, and they rank at number 30 out of 329 sums
+for the largest number of livestock raised per sum (Saizen et al.,
+2010). In the last decade, the number of dust events associated with
+wind erodibility increased by 30 \% in Bayan-Önjüül (Kurosaki et al.,
+2011). This is an area where dust emissions activity has been monitored
+on a long-term basis (Shinoda et al., 2010a) at a dust observation site
+(DOS) adjacent to the study site (Fig. 1a). According to long-term
+meteorological observations made at the monitoring station of the
+Institute of Meteorology and Hydrology of Mongolia located near the
+site, the prevailing wind direction is northwest. Mean annual
+precipitation is 163 mm, and mean temperature is 0.1◦C for the period
+1995 to 2005 (Shinoda et al., 2010b). Soil texture is dominated by sand
+(98.1 \%, with only 1.3 \% clay and 0.6 \% silt; Table 1; Shinoda et
+al., 2010a). Insert figure legends with the first sentence in bold, for
+example:
+
+\newpage
+
+\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_1.png}{]}
+
+\begin{figure}
+\centering
+\includegraphics[width=4.16667in,height=\textheight,keepaspectratio]{images/table_1.png}
+\caption{\textbf{Table 1}. A description of datasets obtained at the
+sites}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/scheme_1.png}
+\caption{Scheme 1. Data handling procedure}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_2b.png}
+\caption{Figure 2. Data gap filling}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_2c.png}
+\caption{Figure 2b. Data gap filling}
+\end{figure}
 

### 2024-11-19 submission/manuscript.Rmd
Drafting Results

@@
 
-Mongolian dust is well-known by its term of coarse, \* color of brown,
-\* its influence on downwind regions, \* and big role in global dust, \*
-and its seasonality.
+## Research Qs
+Therefore, we aimed to demonstrate the distinct temporal and spatial
+variations of PM2.5 and PM10 across urban and rural Mongolia using
+extensive data from 2008 to 2020.
 
-Because, it mainly consists of coarse fractions researchers neglect its
-role in global climate into elaborating global dust-aerosol effects.
-*However, recent change in the driving of the emissions of particulate
-matters might cause fluctuations* its spatio-temporal variations and \*
-characteristics and \* emerge a new pattern, eventually it may lead to
-the role in the regional and global dust, so on climate system through
-\* altering solar incidence, cloud formation, and precipitation or
-warming \* the climate with the fraction of finer particulates.
-
-Mongolia might turn into the not only coarse-dust source region but also
-fine-mixed coarse-dust region.
-
--   Use of coal is increased
-
--   Urbanization
-
--   The monthly mean concentrations of PM10 (PM2.5) reached annual
-    maximum in December and January due to winter synoptic governing
-    conditions in Ulaanbaatar, capital city of Mongolia (Jugder).
-
--   Despite this, the spring dust storms creates another polluted season
-    in UB. On spring, the dust storm from the Gobi Desert contribute
+On spring, the dust storm from the Gobi Desert contribute
 significantly to increased aerosols in the atmosphere and ambient
@@
 $\mu g m^{-3}$ per hour (Jugder).
-
--   Siberian anticyclonic activity governed over Mongolia, which create
-    a significant vulnerability to winter air pollution in the populated
-    areas.
-
--   A such changes in PM10 and PM2.5 to stagnant weather conditions, and
-    local or transported dust was also observed in other countries China
-    (Wang), Korea (Kim) and Japan ().
-
--   deposition
-
--   concentrations of particulate matter is ephederemal, yet vary
+concentrations of particulate matter is ephederemal, yet vary
     depending on whether the pollution cause is natural or industrial,
@@
     challenging.
+    1.  Do concentrations of particulate matters differ in between urban
+        and rural sites, and even within Gobi sites?
+    2.  Do distinct temporal variations has existed among the sites?
+    3.  Do PM2.5 particulates has contributed to the PM10 annual
+        variations?
 
-Therefore, we aimed to demonstrate the distinct temporal and spatial
-variations of PM2.5 and PM10 across urban and rural Mongolia using
-extensive data from 2008 to 2020.
+    -   If yes, how much, and when and where?
+    -   What is the sd, mean, and median
+        -   box plot
+        -   violin
+        -   scatter points, epidemic, sporadic
+    -   Daily variations to examine it related to the heating
+        -   2 peaks: smaller and bigger
+        -   compare the t-duration exceeds 50mug/m3/hour
 
--   The present study will contribute significantly to the understanding
+    4.  Does it has distinct patterns among the sites regarding to the
+        drivings
+
+    -   How PMs varies with the wind speed and visibility
+    -   Do they differently explained with variables and changes in
+        drivings (with PCA analysis)
+
+    5.  Is there any significant changes in time-series of PMs at 4
+        seasons
+    6.  Is there any significant changes in ratio in the spring in
+        respect to winter?
+
+        The present study will contribute significantly to the understanding
             of air particulate matter patterns in Mongolia and providing
             comprehensive data insights for policymakers and public health
-    sectors.
--   Our findings is useful not only for addressing national health
+            sectors. Our findings is useful not only for addressing national health
             impacts but also beneficial for understanding air particulate matter
@@
             the downwind regions in South-east Asia.
+
+## Results
+### The spatio-temporal variations of the PMs at the study sites
+To comparatively examine the spatio-difference regarding to the emissions of PMs, we illustrated the hourly observed values of PM10 and PM2.5 for all sites (figure_3).
+On each site, PM10 is more sporadic than PM2.5 due to dust epidemic nature.
+In conjunction, the mean values averaged from hourly values are larger than its median concentrations both on PM10 and PM2.5 for all sites.
+The mean variables of p-value show that concentrations of particulate matters significantly differ with 99% confident levels at all sites (figure_3), except  95% confident level between DZ and UB on PM10 (figure_3a).
+It clearly exhibits general variations of particulate matters as expected and emphasizes that the significant difference in the concentration values among the sites.
+Besides all sites are quantitatively differ by the values of concentrations in particulate matters, there are two main characteristics can be seen when we compare the median deviations from its mean values etc.
+
+- such patterns are strongly manifested in ZU and SS sites.  
+- the mean values those averaged from hourly concentrations of PM10 and PM2.5 is larger than its median are strongly manifested for all in UB and DZ sites.
+- Additionally, consider the significance of the ..
+
+This highlight that the fine and coarse particulate matters significantly vary among sites of urban and rural sites, and even within Gobi (rural) sites.
+Moreover, .... has a diversely characteristically diversified in respect with the PM2.5 (fine) to PM10 (coarse) particulates.,. which requires urban impacts
+
+To reveal the natural and anthropogenic impacts on the concentrations of particulate matters, we demonstrated annual variations of PM10 and PM2.5 for each sites (figure_4).
+3.  Do PM2.5 particulates has contributed to the PM10 annual
+    variations?
+      Distinct temporal variations has existed among the sites.
+        -   PM2.5 particulates has contributed to the PM10 annual
+            variations in UB and in DZ in winter.
+            -   If yes, how much, and when and where?
+            -   What is the sd, mean, and median
+                -   box plot
+                -   violin
+                -   scatter points, epidemic, sporadic
+            -   Daily variations to examine it related to the heating
+                -   2 peaks: smaller and bigger
+                -   compare the t-duration exceeds 50mug/m3/hour
+
+[Therefore] SS and ZU sites are mainly affected by the spring dust, followed by the autumn dust. Annual maximum in the winter for DZ and UB are from PM2.5, which results an increase in PM10.
+It requires the cause the behind such the variations. DZ site is polluted in the winter by the heating and in the spring by the natural dust.
+
+[Therefore] Spatio-temporally in two class; consists of 2 Gobi sites, and 1 urban plus urbanized Gobi sites.
+
+### The emission patterns of interrelations among meteorological variables at the study sites
+
+        -   PMs varies with the wind speed and visibility
+        -   In general, three distinct patterns were resulted with PCA
+            analysis, which is in consistent with temporal variation.
+            explained with variables and changes in drivings (with PCA
+            analysis)
+
+DZ site is polluted in the winter by the heating and in the spring by the natural dust.
+
+### The recent trends in concentrations of PMS and fine-coarse
+
+        fractional changes at the sites
+        -   There are significant changes in time-series of PMs at 4
+            seasons
+        -   There any significant changes in ratio in the spring in
+            respect to winter in DZ.
+        -   Close relationships was found between PM2.5 in winter and r
+            values in the spring.
+
+## Conclusions
+
+-   The spatio-temporal variations of the PMs at the study sites
+        -   Concentrations of particulate matters differ in between
+            urban and rural sites, and even within Gobi sites.
+        -   Distinct temporal variations has existed among the sites.
+        -   PM2.5 particulates has contributed to the PM10 annual
+            variations in UB and in DZ in winter.
+            -   If yes, how much, and when and where?
+            -   What is the sd, mean, and median
+                -   box plot
+                -   violin
+                -   scatter points, epidemic, sporadic
+            -   Daily variations to examine it related to the heating
+                -   2 peaks: smaller and bigger
+                -   compare the t-duration exceeds 50mug/m3/hour
+    -   The emission patterns of interrelations among meteorological
+        variables at the study sites
+        -   PMs varies with the wind speed and visibility
+        -   In general, three distinct patterns were resulted with PCA
+            analysis, which is in consistent with temporal variation.
+            explained with variables and changes in drivings (with PCA
+            analysis)
+    -   The recent trends in concentrations of PMS and fine-coarse
+        fractional changes at the sites
+        -   There are significant changes in time-series of PMs at 4
+            seasons
+        -   There any significant changes in ratio in the spring in
+            respect to winter PM2.5 in DZ.
+    -   Close relationships was found between PM2.5 in winter and r
+        values in the spring. Thus, our research results clearly proves
+        the distinct variations in PMs has emerged. The dust fine-coarse
+        fractions was manifested at the town center for the Gobi sites,
+        which reveals the that Mongolian dust composites not only coarse
+        dust, but also fine particulate matters. The particulates likely
+        consisted of the black carbon, which may give a substantial
+        effect on climate systems. if this trend continues on as coal
+        consumption with the population growth in the future.
+    -   CO Carbon monoxide is obtained due to incomplete combustion of
+        charcoal in a closed room.
+    -   CO2
 
@@
 1.  Clear annual variations at UB and DZ from pm2.5 pollutions 2. at ZU,
-and SS has a seasonally peaks episodic spring and late autumn from PM10
+    and SS has a seasonally peaks episodic spring and late autumn from
+    PM10
 

### 2024-11-19 submission/manuscript.tex
Drafting Results

@@
 
-Mongolian dust is well-known by its term of coarse, * color of brown, *
-its influence on downwind regions, * and big role in global dust, * and
-its seasonality.
-
-Because, it mainly consists of coarse fractions researchers neglect its
-role in global climate into elaborating global dust-aerosol effects.
-\emph{However, recent change in the driving of the emissions of
-particulate matters might cause fluctuations} its spatio-temporal
-variations and * characteristics and * emerge a new pattern, eventually
-it may lead to the role in the regional and global dust, so on climate
-system through * altering solar incidence, cloud formation, and
-precipitation or warming * the climate with the fraction of finer
-particulates.
-
-Mongolia might turn into the not only coarse-dust source region but also
-fine-mixed coarse-dust region.
-
-\begin{itemize}
-\item
-  Use of coal is increased
-\item
-  Urbanization
-\item
-  The monthly mean concentrations of PM10 (PM2.5) reached annual maximum
-  in December and January due to winter synoptic governing conditions in
-  Ulaanbaatar, capital city of Mongolia (Jugder).
-\item
-  Despite this, the spring dust storms creates another polluted season
-  in UB. On spring, the dust storm from the Gobi Desert contribute
-  significantly to increased aerosols in the atmosphere and ambient air
-  pollution, leading to sporadic peaks in PM10 concentrations reaching
-  as high as 64-234 \(\mu g m^{-3}\) per day or exceeding 6000
-  \(\mu g m^{-3}\) per hour (Jugder).
-\item
-  Siberian anticyclonic activity governed over Mongolia, which create a
-  significant vulnerability to winter air pollution in the populated
-  areas.
-\item
-  A such changes in PM10 and PM2.5 to stagnant weather conditions, and
-  local or transported dust was also observed in other countries China
-  (Wang), Korea (Kim) and Japan ().
-\item
-  deposition
-\item
-  concentrations of particulate matter is ephederemal, yet vary
-  depending on whether the pollution cause is natural or industrial,
-  local or transported, seasonal or non-seasonal, makes complex and
-  challenging.
-\end{itemize}
+\subsection{Research Qs}\label{research-qs}
 
@@
 
+On spring, the dust storm from the Gobi Desert contribute significantly
+to increased aerosols in the atmosphere and ambient air pollution,
+leading to sporadic peaks in PM10 concentrations reaching as high as
+64-234 \(\mu g m^{-3}\) per day or exceeding 6000 \(\mu g m^{-3}\) per
+hour (Jugder). concentrations of particulate matter is ephederemal, yet
+vary depending on whether the pollution cause is natural or industrial,
+local or transported, seasonal or non-seasonal, makes complex and
+challenging. 1. Do concentrations of particulate matters differ in
+between urban and rural sites, and even within Gobi sites? 2. Do
+distinct temporal variations has existed among the sites? 3. Do PM2.5
+particulates has contributed to the PM10 annual variations?
+
+\begin{verbatim}
+-   If yes, how much, and when and where?
+-   What is the sd, mean, and median
+    -   box plot
+    -   violin
+    -   scatter points, epidemic, sporadic
+-   Daily variations to examine it related to the heating
+    -   2 peaks: smaller and bigger
+    -   compare the t-duration exceeds 50mug/m3/hour
+
+4.  Does it has distinct patterns among the sites regarding to the
+    drivings
+
+-   How PMs varies with the wind speed and visibility
+-   Do they differently explained with variables and changes in
+    drivings (with PCA analysis)
+
+5.  Is there any significant changes in time-series of PMs at 4
+    seasons
+6.  Is there any significant changes in ratio in the spring in
+    respect to winter?
+
+    The present study will contribute significantly to the understanding
+        of air particulate matter patterns in Mongolia and providing
+        comprehensive data insights for policymakers and public health
+        sectors. Our findings is useful not only for addressing national health
+        impacts but also beneficial for understanding air particulate matter
+        as ambient air pollution, and tackling atmospheric aerosol effects
+        in the climate system, and revealing their transboundary effects to
+        the downwind regions in South-east Asia.
+\end{verbatim}
+
+\subsection{Results}\label{results}
+
+\subsubsection{The spatio-temporal variations of the PMs at the study
+sites}\label{the-spatio-temporal-variations-of-the-pms-at-the-study-sites}
+
+To comparatively examine the spatio-difference regarding to the
+emissions of PMs, we illustrated the hourly observed values of PM10 and
+PM2.5 for all sites (figure\_3). On each site, PM10 is more sporadic
+than PM2.5 due to dust epidemic nature. In conjunction, the mean values
+averaged from hourly values are larger than its median concentrations
+both on PM10 and PM2.5 for all sites. The mean variables of p-value show
+that concentrations of particulate matters significantly differ with
+99\% confident levels at all sites (figure\_3), except 95\% confident
+level between DZ and UB on PM10 (figure\_3a). It clearly exhibits
+general variations of particulate matters as expected and emphasizes
+that the significant difference in the concentration values among the
+sites. Besides all sites are quantitatively differ by the values of
+concentrations in particulate matters, there are two main
+characteristics can be seen when we compare the median deviations from
+its mean values etc.
+
 \begin{itemize}
@@
 \item
-  The present study will contribute significantly to the understanding
-  of air particulate matter patterns in Mongolia and providing
-  comprehensive data insights for policymakers and public health
-  sectors.
+  such patterns are strongly manifested in ZU and SS sites.\\
 \item
-  Our findings is useful not only for addressing national health impacts
-  but also beneficial for understanding air particulate matter as
-  ambient air pollution, and tackling atmospheric aerosol effects in the
-  climate system, and revealing their transboundary effects to the
-  downwind regions in South-east Asia.
+  the mean values those averaged from hourly concentrations of PM10 and
+  PM2.5 is larger than its median are strongly manifested for all in UB
+  and DZ sites.
+\item
+  Additionally, consider the significance of the ..
+\end{itemize}
+
+This highlight that the fine and coarse particulate matters
+significantly vary among sites of urban and rural sites, and even within
+Gobi (rural) sites. Moreover, \ldots. has a diversely characteristically
+diversified in respect with the PM2.5 (fine) to PM10 (coarse)
+particulates.,. which requires urban impacts
+
+To reveal the natural and anthropogenic impacts on the concentrations of
+particulate matters, we demonstrated annual variations of PM10 and PM2.5
+for each sites (figure\_4). 3. Do PM2.5 particulates has contributed to
+the PM10 annual variations? Distinct temporal variations has existed
+among the sites. - PM2.5 particulates has contributed to the PM10 annual
+variations in UB and in DZ in winter. - If yes, how much, and when and
+where? - What is the sd, mean, and median - box plot - violin - scatter
+points, epidemic, sporadic - Daily variations to examine it related to
+the heating - 2 peaks: smaller and bigger - compare the t-duration
+exceeds 50mug/m3/hour
+
+{[}Therefore{]} SS and ZU sites are mainly affected by the spring dust,
+followed by the autumn dust. Annual maximum in the winter for DZ and UB
+are from PM2.5, which results an increase in PM10. It requires the cause
+the behind such the variations. DZ site is polluted in the winter by the
+heating and in the spring by the natural dust.
+
+{[}Therefore{]} Spatio-temporally in two class; consists of 2 Gobi
+sites, and 1 urban plus urbanized Gobi sites.
+
+\subsubsection{The emission patterns of interrelations among
+meteorological variables at the study
+sites}\label{the-emission-patterns-of-interrelations-among-meteorological-variables-at-the-study-sites}
+
+\begin{verbatim}
+    -   PMs varies with the wind speed and visibility
+    -   In general, three distinct patterns were resulted with PCA
+        analysis, which is in consistent with temporal variation.
+        explained with variables and changes in drivings (with PCA
+        analysis)
+\end{verbatim}
+
+DZ site is polluted in the winter by the heating and in the spring by
+the natural dust.
+
+\subsubsection{The recent trends in concentrations of PMS and
+fine-coarse}\label{the-recent-trends-in-concentrations-of-pms-and-fine-coarse}
+
+\begin{verbatim}
+    fractional changes at the sites
+    -   There are significant changes in time-series of PMs at 4
+        seasons
+    -   There any significant changes in ratio in the spring in
+        respect to winter in DZ.
+    -   Close relationships was found between PM2.5 in winter and r
+        values in the spring.
+\end{verbatim}
+
+\subsection{Conclusions}\label{conclusions}
+
+\begin{itemize}
+\tightlist
+\item
+  The spatio-temporal variations of the PMs at the study sites -
+  Concentrations of particulate matters differ in between urban and
+  rural sites, and even within Gobi sites. - Distinct temporal
+  variations has existed among the sites. - PM2.5 particulates has
+  contributed to the PM10 annual variations in UB and in DZ in winter. -
+  If yes, how much, and when and where? - What is the sd, mean, and
+  median - box plot - violin - scatter points, epidemic, sporadic -
+  Daily variations to examine it related to the heating - 2 peaks:
+  smaller and bigger - compare the t-duration exceeds 50mug/m3/hour
+
+  \begin{itemize}
+  \tightlist
+  \item
+    The emission patterns of interrelations among meteorological
+    variables at the study sites
+
+    \begin{itemize}
+    \tightlist
+    \item
+      PMs varies with the wind speed and visibility
+    \item
+      In general, three distinct patterns were resulted with PCA
+      analysis, which is in consistent with temporal variation.
+      explained with variables and changes in drivings (with PCA
+      analysis)
+    \end{itemize}
+  \item
+    The recent trends in concentrations of PMS and fine-coarse
+    fractional changes at the sites
+
+    \begin{itemize}
+    \tightlist
+    \item
+      There are significant changes in time-series of PMs at 4 seasons
+    \item
+      There any significant changes in ratio in the spring in respect to
+      winter PM2.5 in DZ.
+    \end{itemize}
+  \item
+    Close relationships was found between PM2.5 in winter and r values
+    in the spring. Thus, our research results clearly proves the
+    distinct variations in PMs has emerged. The dust fine-coarse
+    fractions was manifested at the town center for the Gobi sites,
+    which reveals the that Mongolian dust composites not only coarse
+    dust, but also fine particulate matters. The particulates likely
+    consisted of the black carbon, which may give a substantial effect
+    on climate systems. if this trend continues on as coal consumption
+    with the population growth in the future.
+  \item
+    CO Carbon monoxide is obtained due to incomplete combustion of
+    charcoal in a closed room.
+  \item
+    CO2
+  \end{itemize}
 \end{itemize}
@@
 
-\subsection{Conclusions}\label{conclusions}
+\subsection{Conclusions}\label{conclusions-1}
 

### 2024-11-19 submission/manuscript.Rmd
MS draft has started

@@
 
-## Abstract
+## Abstract (150 words)
 
@@
 ## Introduction
+    * Advanced the knowledge of global dust, has reached to recognize the sources,.
+      - Classification dust brown color, seasonal characteristics, with coarse fractions.
+      - This knowledge further efficient to climate system when elaborating dust-aerosol effects.
+      - But, a large uncertainties in the global dust model has existed so for climate models which clearly limits our understanding the climate system and shape the facing global issues of global warming.
+      - This is mainly caused by the lack of parameterization and recognition of iterative changes controlled by the natural forces and anthropogenic drivings.
+
+    * Mongolian dust brown color, seasonal characteristics, with coarse fractions.
+      - Mongolian dust has an attention of the its mass fraction in global dust, yet unlikely elaborated in the climate models due to its majority of coarse fraction for its a small contribution to the climate system through its radiative feedback.
+      - But, such recognized characterization might get no longer valid due to recent change in the driving of the emissions of air particulate matters.
+A large high concentrations of PM2.5 in the capital city of Mongolia has been observed  as a result of the heavy consumptions of coal as a winter heating has rapidly spread as a mining industry taken off since 2000.
+Winter weather stagnant conditions governed by the Siberian magnifies the concentrations of the particulate matter emissions by trapping the polluted air below the boundary layer, so that results in a very large high concentrations of PM2.5, locally.
+Even  recognized as one of the highly polluted capital cities in the world.
+      - Therefore, It is important to examine the emerging changes and shifting patterns of air particulate matters in Mongolia.
+More importantly, it is essential to reveal the significant changes in the the altered fraction particularly, in the dust seasons considering its high potential of intriuging in the free atmosphere to transported in the long-distance, so carrying capacity of the role to shift the global climate system, and its side impacts on downwind regions.
+
+* Study goal
+      - We hypothesize ...
+      - Our study will benefit not only to the global dust research but also climate, and further to the country itself for urban planning, and coal combustion.
 
@@
 
+### Materials
+
+### Methods  3,000 words
+
+## Acknowledgements
+Keep acknowledgements brief and do not include thanks to anonymous referees or editors, or effusive
+comments. Grant or contribution numbers may be acknowledged.
+
+## Figures (10)
+Figure legends should be <350 words each. They should begin with a
+brief title sentence for the whole figure and continue with a short
+statement of what is depicted in the figure, not the results (or data) of
+the experiment or the methods used. Legends should be detailed enough
+so that each figure and caption can, as far as possible, be understood in
+isolation from the main text.
+
+Tables. Each table should be prepared using the Table menu in Word or
+the table environment in TeX/LaTeX and accompanied by a short title
+sentence describing what the table shows. Further details can be
+included as footnotes to the table.
+
+## References (70)
+
+## Supplementary
+Author contributions. You must include a statement that specifies the
+individual contributions of each co-author. For example: "A.P.M.
+‘contributed’ Y and Z; B.T.R. ‘contributed’ Y,” etc. See our authorship
+policies for more details.
+
+Competing interests. Submission of a competing interests statement is
+required for all content of the journal.
+
+Materials & Correspondence. Indicate the author(s) to whom
+correspondence and material requests should be addressed.
+
+Supplementary information
+Please submit supplementary figures, small tables and text as a single
+combined PDF document. Tables longer than one page should be
+provided as an Excel or similar file type. For optimal quality video files
+please use H.264 encoding, the standard aspect ratio of 16:9 (4:3 is
+second best) and do not compress the video. We encourage submission
+of step-by-step synthesis procedures for chemical compounds and data
+on compound characterization. Supplementary information is not copyedited, so please ensure that it is clearly and succinctly presented, and
+that the style and terminology conform to the rest of the manuscript.
+
+## Materials and Methods
+
 ### A description of study sites

### 2024-11-22 submission/Outline_20241121.Rmd
draft processing

--- 2024-11-15 submission/manuscript.Rmd
@@
     df_print: paged
-csl: "nature-communications.csl"
+csl: "mbio.csl"
 fontsize: 11pt
@@
 
-Munkh\^1, Joeseph P. Schmo\^2, Sally J. Rivers\^1, Patrick D.
-Schloss^1^$\dagger$
+Munkhtsetseg, Shimizu. A, Matsuki. A, Batdorj. D, Matsui. H
 
@@
 
-$\dagger$ To whom correspondence should be addressed:
-[pschloss\@umich.edu](mailto:pschloss@umich.edu){.email}
-
-1\. Department of Microbiology and Immunology, University of Michigan,
-Ann Arbor, MI 48109
-
-2\. Other department contact information
+To whom correspondence should be addressed:
 
@@
 7.  r ratio shows ... emission source; dust might carry anthropogenic
-    fine particulates as well. \newpage
+    fine particulates as well.
+
+\newpage
 
@@
 
-Mongolian dust is well-known by its term of coarse, \* color of brown,
-\* its influence on downwind regions, \* and big role in global dust, \*
-and its seasonality.
+-   Advanced the knowledge of global dust, has reached to recognize the
+    sources,.
 
-Because, it mainly consists of coarse fractions researchers neglect its
-role in global climate into elaborating global dust-aerosol effects.
-*However, recent change in the driving of the emissions of particulate
-matters might cause fluctuations* its spatio-temporal variations and \*
-characteristics and \* emerge a new pattern, eventually it may lead to
-the role in the regional and global dust, so on climate system through
-\* altering solar incidence, cloud formation, and precipitation or
-warming \* the climate with the fraction of finer particulates.
+-   Classification dust brown color, seasonal characteristics, with
+    coarse fractions.
 
-Mongolia might turn into the not only coarse-dust source region but also
-fine-mixed coarse-dust region.
+-   This knowledge further efficient to climate system when elaborating
+    dust-aerosol effects.
 
--   Use of coal is increased
+-   But, a large uncertainties in the global dust model has existed so
+    for climate models which clearly limits our understanding the
+    climate system and shape the facing global issues of global warming.
 
--   Urbanization
+-   This is mainly caused by the lack of parameterization and
+    recognition of iterative changes controlled by the natural forces
+    and anthropogenic drivings.
 
--   The monthly mean concentrations of PM10 (PM2.5) reached annual
-    maximum in December and January due to winter synoptic governing
-    conditions in Ulaanbaatar, capital city of Mongolia (Jugder).
+-   Mongolian dust brown color, seasonal characteristics, with coarse
+    fractions.
 
--   Despite this, the spring dust storms creates another polluted season
-    in UB. On spring, the dust storm from the Gobi Desert contribute
-    significantly to increased aerosols in the atmosphere and ambient
-    air pollution, leading to sporadic peaks in PM10 concentrations
-    reaching as high as 64-234 $\mu g m^{-3}$ per day or exceeding 6000
-    $\mu g m^{-3}$ per hour (Jugder).
+-   Mongolian dust has an attention of the its mass fraction in global
+    dust, yet unlikely elaborated in the climate models due to its
+    majority of coarse fraction for its a small contribution to the
+    climate system through its radiative feedback.
 
--   Siberian anticyclonic activity governed over Mongolia, which create
-    a significant vulnerability to winter air pollution in the populated
-    areas.
+-   But, such recognized characterization might get no longer valid due
+    to recent change in the driving of the emissions of air particulate
+    matters. A large high concentrations of PM2.5 in the capital city of
+    Mongolia has been observed as a result of the heavy consumptions of
+    coal as a winter heating has rapidly spread as a mining industry
+    taken off since 2000. Winter weather stagnant conditions governed by
+    the Siberian magnifies the concentrations of the particulate matter
+    emissions by trapping the polluted air below the boundary layer, so
+    that results in a very large high concentrations of PM2.5, locally.
+    Even recognized as one of the highly polluted capital cities in the
+    world.
 
--   A such changes in PM10 and PM2.5 to stagnant weather conditions, and
-    local or transported dust was also observed in other countries China
-    (Wang), Korea (Kim) and Japan ().
+-   Therefore, It is important to examine the emerging changes and
+    shifting patterns of air particulate matters in Mongolia. More
+    importantly, it is essential to reveal the significant changes in
+    the the altered fraction particularly, in the dust seasons
+    considering its high potential of intriuging in the free atmosphere
+    to transported in the long-distance, so carrying capacity of the
+    role to shift the global climate system, and its side impacts on
+    downwind regions.
 
--   deposition
+-   Study goal - We hypothesize ... - Our study will benefit not only to
+    the global dust research but also climate, and further to the
+    country itself for urban planning, and coal combustion.
 
--   concentrations of particulate matter is ephederemal, yet vary
-    depending on whether the pollution cause is natural or industrial,
-    local or transported, seasonal or non-seasonal, makes complex and
-    challenging.
+### Research Qs
 
@@
 
--   The present study will contribute significantly to the understanding
-    of air particulate matter patterns in Mongolia and providing
-    comprehensive data insights for policymakers and public health
-    sectors.
--   Our findings is useful not only for addressing national health
-    impacts but also beneficial for understanding air particulate matter
-    as ambient air pollution, and tackling atmospheric aerosol effects
-    in the climate system, and revealing their transboundary effects to
-    the downwind regions in South-east Asia.
+On spring, the dust storm from the Gobi Desert contribute significantly
+to increased aerosols in the atmosphere and ambient air pollution,
+leading to sporadic peaks in PM10 concentrations reaching as high as
+64-234 $\mu g m^{-3}$ per day or exceeding 6000 $\mu g m^{-3}$ per hour
+(Jugder). concentrations of particulate matter is ephederemal, yet vary
+depending on whether the pollution cause is natural or industrial, local
+or transported, seasonal or non-seasonal, makes complex and
+challenging. 1. Do concentrations of particulate matters differ in
+between urban and rural sites, and even within Gobi sites? 2. Do
+distinct temporal variations has existed among the sites? 3. Do PM2.5
+particulates has contributed to the PM10 annual variations?
+
+-   If yes, how much, and when and where?
+-   What is the sd, mean, and median
+    -   box plot
+    -   violin
+    -   scatter points, epidemic, sporadic
+-   Daily variations to examine it related to the heating
+    -   2 peaks: smaller and bigger
+    -   compare the t-duration exceeds 50mug/m3/hour
+
+4.  Does it has distinct patterns among the sites regarding to the
+    drivings
+
+-   How PMs varies with the wind speed and visibility
+-   Do they differently explained with variables and changes in drivings
+    (with PCA analysis)
+
+5.  Is there any significant changes in time-series of PMs at 4 seasons
+6.  Is there any significant changes in ratio in the spring in respect
+    to winter?
+
+The present study will contribute significantly to the understanding of
+air particulate matter patterns in Mongolia and providing comprehensive
+data insights for policymakers and public health sectors. Our findings
+is useful not only for addressing national health impacts but also
+beneficial for understanding air particulate matter as ambient air
+pollution, and tackling atmospheric aerosol effects in the climate
+system, and revealing their transboundary effects to the downwind
+regions in South-east Asia.
+
+## Results
+### The spatio-temporal variations of the PMs at the study sites
+To evaluate the spatial variations in particulate matter (PM) concentrations, we displayed hourly observed values of PM10 and PM2.5 for all study sites (figure_3).
+The mean p-values indicate that PM concentrations differ significantly at a 99% confidence level across all sites (figure_3), with the exception of a 95% confidence level between DZ and UB for PM10 (figure_3a), highlighting substantial concentration disparities among sites.
+While quantitative differences in PM concentration values exist across all sites, two key patterns emerge when examining median deviations from mean values and irregular observation fluctuations.
+For instance, PM10 demonstrates more erratic behavior than PM2.5 at each location, particularly evident at ZU and SS sites.
+Furthermore, the mean values calculated from hourly measurements surpass the median concentrations for both PM10 and PM2.5 across all sites, with notable prominence at UB and DZ locations.
+Consequently, significant spatial differences in PM concentrations exist among all sites, regardless of urban or rural classification.
+However, the sites can be categorized into two groups based on their characteristics: UB (urban) and DZ (rural town, Gobi); and SS (rural, Gobi) and ZU (rural, Gobi).
+These findings for DZ appear to support our hypothesis of emerging new emission patterns related to increased coal consumption during winter months.
+
+--> 3.1.2
+
+To investigate whether emerging PM patterns are associated with household winter heating activities, we demonstrated annual variations in PM10 and PM2.5 concentrations at the sites.
+Significant annual variations in PM10 and PM2.5 levels were observed at UB and DZ sites, with maximum concentrations exceeding 100 $\mu g/m^3$ during colder months (January, November, December) and lower levels consistently below 50 $\mu g/m^3$ during warmer months (May-September).
+These annual maximums coincided with the diurnal variations in PM10 and PM2.5 concentrations at sites DZ and UB, where PM concentrations reached their highest values during nighttime and early morning hours (approximately 8 PM to 4 AM UTC), with median values surpassing 50 $\mu g/m^3$.
+Conversely, both pollutants exhibited reduced concentrations during daytime hours (8AM to 4PM UTC), likely due to increased atmospheric dispersion.
+UB site exhibited similar daily fluctuations with extended periods (approximately 8 PM to 5 AM UTC) of elevated concentrations.
+Additionally, winter PM10 concentrations at both sites were primarily composed of PM2.5 (figure_5; mean values for PM10 with the color bar).
+The increase in PM10 and PM2.5 aligns with the heating active hours, suggesting that household coal consumption contributes to elevated PM levels at both DZ and UB sites.
+In contrast, ZU and SS sites displayed significantly lower annual PM10 and PM2.5 levels, with sudden frequent spikes in spring followed by occasional instances in autumn.
+These annual variations were more pronounced in PM10 compared to PM2.5, highlighting the impact of Gobi dust and sand storms.
+Similar occurrences were also noted at the DZ site, suggesting its exposure to both winter heating emissions and natural spring dust, reflecting its Gobi-region characteristics.
+The annual variability in PMs with higher concentrations during nighttime and colder months, indicating the influence of localized emission sources and reduced boundary layer mixing at UB and DZ sites.
+Additionally, the upward extended ranges without the bottom bottle of the violin plot, demonstrating greater variability during colder months.
+It implies instability in concentrations potentially reaching high levels above 400 $\mu g/m^3$ when Arctic oscillation/Siberian high intensifies with heating, and dropping below 50 $\mu g/m^3$ when it weakens or heating demand decreases.
+These findings confirm the (DZ) emerging PM pattern is caused by household activities and influenced by meteorological conditions.
+Overall, meteorological factors appear to play a crucial role in governing PM levels.
+
+--> 3.2.1
+
+### The emission patterns of interrelations among meteorological variables at the study sites
+  - PMs varies with the wind speed and visibility
+  - In general, three distinct patterns were resulted with PCA analysis, which is in consistent with temporal variation. explained with variables and changes in drivings (with PCA analysis)
+
+[AND] To distinguish the emission driving variables, we demonstrated interrelations between(WS) and visibility (VIS) among particulate matters of PM10 and PM2.5 for study sites (figure_4).
+[AND] Higher PM concentrations tend to correspond to lower wind speeds (smaller circles) and reduced visibility (darker blue points) in UB.
+[AND] All Gobi sites show the largest variability in PM concentrations, with extreme outliers correlated with the WS primarily in the springs.
+
+[AND] In comparison between PM10 increase with the PM2.5, as concentrations increase in PM10 concentration the ratio 1:1 is inscreased approximately 1.3 times for UB, and 2 times for all three Gobi sites.
+This increase in Gobi sites were further increased upto 5 times in ZU, and XXX times in SS sites showing the insufficient PM2.5 particulates in the those sites.
+
+The Principal Component Analysis (PCA) biplot (Figure 7a) highlights the relationships among key variables, including PM10, PM2.5, wind speed (WS), wind direction (WD), visibility, relative humidity (r), and optical properties of aerosols (OPC).
+The first two principal components (Dim 1 and Dim 2) explain 66.52% of the total variance, with Dim 1 (44.36%) strongly associated with PM10 and PM2.5, indicating that these pollutants are major contributors to air quality variations.
+Visibility is inversely correlated with PM10 and PM2.5, suggesting that reduced visibility aligns with higher pollution levels.
+Wind speed and wind direction are positively aligned with Dim 2 (22.16%), reflecting their influence on pollutant dispersion.
+
+The clustering of monitoring sites based on PM concentrations and geographic characteristics (Figure 7b) reveals distinct patterns.
+UB (urban background) is characterized by high PM concentrations (positive Dim 1 scores), while SS (suburban site) demonstrates low PM levels, clustering tightly along the negative Dim 1 axis.
+DZ (industrial site) exhibits substantial variability, with clusters extending into higher Dim 1 and Dim 2 ranges, reflecting the complex interplay of emission sources and meteorological factors.
+ZU (rural site) overlaps with SS but shows greater spread, indicating moderate pollution levels influenced by seasonal and localized factors.
+These results underscore the significant spatial and meteorological influences on air quality across the studied locations.
+
+[BUT]
+
+[THEREFORE] 3 clusters, supply is different?
+
+###  The recent trends in concentrations of PMS and fine-coarse fractional changes at the sites
+Figure 8 illustrates the trend analysis of PM10 and PM2.5 concentrations across four monitoring sites (UB, DZ, ZU, and SS) from 2009 to 2020.
+The time series demonstrates seasonal patterns and trends, with notable differences in particulate matter concentrations between the four seasons: Q1 (winter), Q2 (spring), Q3 (summer), and Q4 (autumn).
+
+The p-values displayed on the figure indicate the statistical significance of trend changes in PM10 and PM2.5 levels for each season.
+At UB, significant negative trends are observed for Q1 and Q2 (p < 0.001).
+At DZ, a significant positive trend is observed for Q1 (R = 0.17, p < 0.001), indicating increasing particulate matter levels during this season.17, p < 0.001), while Q2 exhibits no significant trend (R = 0, p = 0.915)., highlighting a decreasing pattern in particulate concentrations during these periods over time.
+In contrast, Q3 and Q4 show weaker or nonsignificant trends (e.g., R = -0.03, p > 0.2), suggesting minimal changes in PM levels during the warmer months.
+
+The industrial site (DZ) exhibits higher variability in PM levels, with periodic spikes indicative of localized emissions and industrial activities.
+Urban background (UB) consistently records elevated PM levels during Q1, driven by heating and combustion-related sources.
+Rural (ZU) and suburban (SS) sites generally display lower particulate concentrations.
+However, at ZU, a significant negative trend is observed in both PM10 and PM2.5 for Q1, suggesting a consistent reduction in particulate matter during winter months (may reduced transboundary traffic).
+At SS, a negative trend is observed in PM2.5 for Q2, reflecting a seasonal decline potentially linked to  reduced wind speed during spring.
+
+The significant decreasing trends in PM concentrations in UB during colder months likely indicate improvements in emission control measures and a transition to more efficient heating practices.
+In contrast, a significant increasing trend is observed at DZ, particularly during Q1, suggesting rising emissions potentially associated with urbanized  household practices and other local activities in Mongolia's regional towns.
+
+  - There are significant changes in time-series of PMs at 4 seasons
+  - There any significant changes in ratio in the spring in respect to winter in DZ.
+  - Close relationships was found between PM2.5 in winter and r values in the spring.
+
+## Conclusions
+
+In this study, we investigated the temporal variations of PM2.5 and PM10
+concentrations at the 4 sites of rural and urban those located along the
+the wind corridor. Three distinct variations has been detected.
+
+Air quality in urban sites is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
+Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
+Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
+
+A clear seasonal variations in the sites of UB and DZ is [Air quality is
+governed by natural dust emission, and anthropogenic emissions]
+
+* Due
+to rapid increase in urban, and combustion of coal/oyutolgoi for heating
+winter conditions results a highly increase in not only capital city but
+also towns
+* In a result, spring coarse dust, plus winter fine
+pollutants [spring coarse dust is immediately transported and deposited
+in the source area, whereas winter fine pollutants is permanently stayed
+in the source area due to stagnant atmosphere govern over entire
+country., perhaps float- ing in the near surface, deposits in the
+surface]
+* Alarms, the Mongolian dust in the spring, optical properties
+will be shifted; this gives ... Gobi dust and sand storms has become
+tuiren, from the shoroon shuurga. which clearly requires the attention.
+
+Following problems
+
+-   On downwind regions
+-   On national-level Demonstrating temporal and spatial variations of
 
 \newpage
-
-## Results and Discussion
-
-\newpage
-\subsection{Urban and rural distinct Spatio-temporal diverse variations of $PM_{10}$ and $PM_{2.5}$}
 
@@
 
-## Conclusions
-
-In this study, we investigated the temporal variations of PM2.5 and PM10
-concentrations at the 4 sites of rural and urban those located along the
-the wind corridor. Three distinct variations has been detected.
-
-\begin{enumerate}
-    \item Air quality in urban sites is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
-    \item Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
-    \item Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
-\end{enumerate}
-
-A clear seasonal variations in the sites of UB and DZ is [Air quality is
-governed by natural dust emission, and anthropogenic emissions] \* Due
-to rapid increase in urban, and combustion of coal/oyutolgoi for heating
-winter conditions results a highly increase in not only capital city but
-also towns \* In a result, spring coarse dust, plus winter fine
-pollutants [spring coarse dust is immediately transported and deposited
-in the source area, whereas winter fine pollutants is permanently stayed
-in the source area due to stagnant atmosphere govern over entire
-country., perhaps float- ing in the near surface, deposits in the
-surface] \* Alarms, the Mongolian dust in the spring, optical properties
-will be shifted; this gives ... Gobi dust and sand storms has become
-tuiren, from the shoroon shuurga. which clearly requires the attention.
-
-Following problems
-
--   On downwind regions
--   On national-level Demonstrating temporal and spatial variations of
-    air particulate matter has become important for understanding
-    characteristics of particulate matter in the climate system,
-    providing valuable information for well-established air quality
-    measures, and illustrating the good trace data for health studies.
-    Because particulate pollutants have a great impact on human health
-    (Dockery and Pope,1994; Harrison and Yin, 2000; Hong et al., 2002),
-    high atmospheric concentrations of these pollutants was a major
-    concern particularly in urban areas, in the last 2-3 decades. Recent
-    studies highlight that even low concentrations of these pollutants
-    can lead to various health issues, and may associate with morbidity
-    and mortality across the life span (Zigler et al., 2017). Children
-    exposed to high levels of air pollution show increased rates of
-    asthma, decreased lung function growth, and increased risk of early
-    markers of cardiovascular disease (Bourdrel et al., 2017; Gauderman
-    et al., 2015; Hehua et al., 2017). Short-term exposure with high
-    level of PM10 resulted the chronic cardiovascular disease in
-    Mongolia (Enkhjargal 2020). In addition to these health issues,
-    (prenatal) neurodevelopmental impacts such as effects on
-    intelligence, attention, autism, and mood, while aging populations
-    experience accelerated cognitive decline when exposed to high levels
-    of pollution is detected (Power et al., 2016). Long-term exposure to
-    low levels of particulate matter, such as concentrations as low as
-    10 $\mu g m^{-3}$ (equilibrium to WHO Air Quality Guidelines), has
-    been linked to increased lung cancer in the EU (Hvidtfeldt et al.
-    2021), with similar evidences reported in Canada (Bai et al., 2019),
-    and significantly higher rates captured in China with concentrations
-    up to 30 $\mu g m^{-3}$. Apparently, pollutants of particulate
-    matters has effects to various health issues with the different
-    thresholds and exposure durations. However, more in-depth and
-    diversified research on air pollution and its health effects is
-    essential, with the detailed information is necessary (Tan et al
-    2021) to have accuracy of assessing exposure to air pollution during
-          developmentally relevant time periods, such as trimesters or
-          months (Becerra et al., 2013; Gong et al., 2014; Kalkbrenner
-          et al., 2014) or weeks (Chiu et al., 2016). Many research
-          findings/Numerous research findings have advanced the field,
-          and air quality indices is widely used for providing guidance,
-          and public perception of air quality has been improved
-          (Mirabelli et al., 2020).
-
-\newpage
-
-## Materials and Methods
-
-### A description of study sites
-
-According to the spatial magnitude of wind stress in Mongolia (Figure
-1), the largest magnitude of wind speed is on the Gobi sites,
-particularly those located in the southeast edge of the country.
-
--   The impact of high winds on plant diversity varies across
-    environmental gradients of precipitation and soil fertility
-    (Milchunas et al., 1988).
--   In the desert steppe zone, species richness was lower in the drier
-    years but did not vary with grazing pressure.
--   In the steppe zone, species richness varied significantly with
-    grazing pressure but did not vary between years. Species richness is
-    not impacted by grazing gradient in desert steppe, but it is in the
-    steppe (Cheng et al., 2011).
-
-In the last 2 decades, due to poverty and natural disasters there is
-population immigration has taken place from the rural to urban,
-especially to capital city of Mongolia. Due to tiny infrastructure to
-provide the mega city with the dense population, it introduces the urban
-pollution. Therefore, Ulaanbaatar air particulate matter mainly reflects
-the coal burning, and partly, natural dust.
-
-Consequently, the atmospheric environment and climate for Mongolian Gobi
-has been impacted the most by frequent dust and and sand storm in the
-spring.
-
-Our study was carried out in Dalanzadgad (town center) (Tbl. 1; 43.57°N,
-104.42°E), Sainshand (Tbl. 1; 44.87°N, 110.12°E) and Zamyn-Uud (Tbl. 1;
-43.72°N, 111.90°E) in the Gobi Desert, and at Ulaanbaatar (Tbl.??.??°N,
-104.42°E) (city center) located in the temperate Mongolian steppe of
-Mongolia (Figure 2). Nomads and settlements of this sum have raised a
-large number of livestock, and they rank at number 30 out of 329 sums
-for the largest number of livestock raised per sum (Saizen et al.,
-2010). In the last decade, the number of dust events associated with
-wind erodibility increased by 30 % in Bayan-Önjüül (Kurosaki et al.,
-2011). This is an area where dust emissions activity has been monitored
-on a long-term basis (Shinoda et al., 2010a) at a dust observation site
-(DOS) adjacent to the study site (Fig. 1a). According to long-term
-meteorological observations made at the monitoring station of the
-Institute of Meteorology and Hydrology of Mongolia located near the
-site, the prevailing wind direction is northwest. Mean annual
-precipitation is 163 mm, and mean temperature is 0.1◦C for the period
-1995 to 2005 (Shinoda et al., 2010b). Soil texture is dominated by sand
-(98.1 %, with only 1.3 % clay and 0.6 % silt; Table 1; Shinoda et al.,
-2010a). Insert figure legends with the first sentence in bold, for
-example:
-
-\newpage
-
 ![Geographic locations of study
@@
 
+[@Munkh2017]
+
 \newpage

### 2024-11-22 submission/Outline_20241121.tex
draft processing

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  11pt,
]{article}
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}

\usepackage[none]{hyphenat}

\usepackage{setspace}
\doublespacing
\setlength{\parskip}{1em}

\usepackage{lineno}

\usepackage{pdfpages}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Spatio-temporal distinct patterns in variations of PM\_\{10\} and PM\_\{2.5\} relative to the recent drivings of emission sources in Mongolia},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\textbf{Spatio-temporal distinct patterns in variations of
\(PM_{10}\) and \(PM_{2.5}\) relative to the recent drivings of emission
sources in Mongolia}}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

\vspace{35mm}

Running title: INSERT RUNNING TITLE HERE

\vspace{35mm}

Munkhtsetseg, Shimizu. A, Matsuki. A, Batdorj. D, Matsui. H

\vspace{40mm}

To whom correspondence should be addressed:

\newpage
\linenumbers

\subsection{Abstract (150 words)}\label{abstract-150-words}

Storyline:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  A new pattern is emerged
\item
  Air quality in urban sites is episodically dictated by dust events in
  spring or late autumn, yet seasonally governed by anthropogenic
  emissions in winter. {[}Air quality is governed by natural dust
  emission, and anthropogenic emissions{]}
\item
  With recent growing interest in urban life style, and combustion of
  coal/oyutolgoi for heating winter conditions results a highly increase
  in not only capital city but also towns
\item
  In a result, spring coarse dust, plus winter fine pollutants
\item
  spring coarse dust is immediately transported and deposited in the
  source area, whereas winter fine pollutants is permanently stayed in
  the source area due to stagnant atmosphere govern over entire
  country., perhaps floating in the near surface, deposits in the
  surface{]}
\item
  Alarms, the Mongolian dust in the spring, optical properties might be
  shifted; this gives \ldots{} Gobi dust and sand storms has become
  tuiren, from the shoroon shuurga. which clearly requires the
  attention.
\item
  r ratio shows \ldots{} emission source; dust might carry anthropogenic
  fine particulates as well.
\end{enumerate}

\newpage

\subsection{Introduction}\label{introduction}

\begin{itemize}
\item
  Advanced the knowledge of global dust, has reached to recognize the
  sources,.
\item
  Classification dust brown color, seasonal characteristics, with coarse
  fractions.
\item
  This knowledge further efficient to climate system when elaborating
  dust-aerosol effects.
\item
  But, a large uncertainties in the global dust model has existed so for
  climate models which clearly limits our understanding the climate
  system and shape the facing global issues of global warming.
\item
  This is mainly caused by the lack of parameterization and recognition
  of iterative changes controlled by the natural forces and
  anthropogenic drivings.
\item
  Mongolian dust brown color, seasonal characteristics, with coarse
  fractions.
\item
  Mongolian dust has an attention of the its mass fraction in global
  dust, yet unlikely elaborated in the climate models due to its
  majority of coarse fraction for its a small contribution to the
  climate system through its radiative feedback.
\item
  But, such recognized characterization might get no longer valid due to
  recent change in the driving of the emissions of air particulate
  matters. A large high concentrations of PM2.5 in the capital city of
  Mongolia has been observed as a result of the heavy consumptions of
  coal as a winter heating has rapidly spread as a mining industry taken
  off since 2000. Winter weather stagnant conditions governed by the
  Siberian magnifies the concentrations of the particulate matter
  emissions by trapping the polluted air below the boundary layer, so
  that results in a very large high concentrations of PM2.5, locally.
  Even recognized as one of the highly polluted capital cities in the
  world.
\item
  Therefore, It is important to examine the emerging changes and
  shifting patterns of air particulate matters in Mongolia. More
  importantly, it is essential to reveal the significant changes in the
  the altered fraction particularly, in the dust seasons considering its
  high potential of intriuging in the free atmosphere to transported in
  the long-distance, so carrying capacity of the role to shift the
  global climate system, and its side impacts on downwind regions.
\item
  Study goal - We hypothesize \ldots{} - Our study will benefit not only
  to the global dust research but also climate, and further to the
  country itself for urban planning, and coal combustion.
\end{itemize}

\subsubsection{Research Qs}\label{research-qs}

Therefore, we aimed to demonstrate the distinct temporal and spatial
variations of PM2.5 and PM10 across urban and rural Mongolia using
extensive data from 2008 to 2020.

On spring, the dust storm from the Gobi Desert contribute significantly
to increased aerosols in the atmosphere and ambient air pollution,
leading to sporadic peaks in PM10 concentrations reaching as high as
64-234 \(\mu g m^{-3}\) per day or exceeding 6000 \(\mu g m^{-3}\) per
hour (Jugder). concentrations of particulate matter is ephederemal, yet
vary depending on whether the pollution cause is natural or industrial,
local or transported, seasonal or non-seasonal, makes complex and
challenging. 1. Do concentrations of particulate matters differ in
between urban and rural sites, and even within Gobi sites? 2. Do
distinct temporal variations has existed among the sites? 3. Do PM2.5
particulates has contributed to the PM10 annual variations?

\begin{itemize}
\tightlist
\item
  If yes, how much, and when and where?
\item
  What is the sd, mean, and median

  \begin{itemize}
  \tightlist
  \item
    box plot
  \item
    violin
  \item
    scatter points, epidemic, sporadic
  \end{itemize}
\item
  Daily variations to examine it related to the heating

  \begin{itemize}
  \tightlist
  \item
    2 peaks: smaller and bigger
  \item
    compare the t-duration exceeds 50mug/m3/hour
  \end{itemize}
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Does it has distinct patterns among the sites regarding to the
  drivings
\end{enumerate}

\begin{itemize}
\tightlist
\item
  How PMs varies with the wind speed and visibility
\item
  Do they differently explained with variables and changes in drivings
  (with PCA analysis)
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  Is there any significant changes in time-series of PMs at 4 seasons
\item
  Is there any significant changes in ratio in the spring in respect to
  winter?
\end{enumerate}

The present study will contribute significantly to the understanding of
air particulate matter patterns in Mongolia and providing comprehensive
data insights for policymakers and public health sectors. Our findings
is useful not only for addressing national health impacts but also
beneficial for understanding air particulate matter as ambient air
pollution, and tackling atmospheric aerosol effects in the climate
system, and revealing their transboundary effects to the downwind
regions in South-east Asia.

\subsection{Results}\label{results}

\subsubsection{The spatio-temporal variations of the PMs at the study
sites}\label{the-spatio-temporal-variations-of-the-pms-at-the-study-sites}

To evaluate the spatial variations in particulate matter (PM)
concentrations, we displayed hourly observed values of PM10 and PM2.5
for all study sites (figure\_3). The mean p-values indicate that PM
concentrations differ significantly at a 99\% confidence level across
all sites (figure\_3), with the exception of a 95\% confidence level
between DZ and UB for PM10 (figure\_3a), highlighting substantial
concentration disparities among sites. While quantitative differences in
PM concentration values exist across all sites, two key patterns emerge
when examining median deviations from mean values and irregular
observation fluctuations. For instance, PM10 demonstrates more erratic
behavior than PM2.5 at each location, particularly evident at ZU and SS
sites. Furthermore, the mean values calculated from hourly measurements
surpass the median concentrations for both PM10 and PM2.5 across all
sites, with notable prominence at UB and DZ locations. Consequently,
significant spatial differences in PM concentrations exist among all
sites, regardless of urban or rural classification. However, the sites
can be categorized into two groups based on their characteristics: UB
(urban) and DZ (rural town, Gobi); and SS (rural, Gobi) and ZU (rural,
Gobi). These findings for DZ appear to support our hypothesis of
emerging new emission patterns related to increased coal consumption
during winter months.

--\textgreater{} 3.1.2

To investigate whether emerging PM patterns are associated with
household winter heating activities, we demonstrated annual variations
in PM10 and PM2.5 concentrations at the sites. Significant annual
variations in PM10 and PM2.5 levels were observed at UB and DZ sites,
with maximum concentrations exceeding 100 \(\mu g/m^3\) during colder
months (January, November, December) and lower levels consistently below
50 \(\mu g/m^3\) during warmer months (May-September). These annual
maximums coincided with the diurnal variations in PM10 and PM2.5
concentrations at sites DZ and UB, where PM concentrations reached their
highest values during nighttime and early morning hours (approximately 8
PM to 4 AM UTC), with median values surpassing 50 \(\mu g/m^3\).
Conversely, both pollutants exhibited reduced concentrations during
daytime hours (8AM to 4PM UTC), likely due to increased atmospheric
dispersion. UB site exhibited similar daily fluctuations with extended
periods (approximately 8 PM to 5 AM UTC) of elevated concentrations.
Additionally, winter PM10 concentrations at both sites were primarily
composed of PM2.5 (figure\_5; mean values for PM10 with the color bar).
The increase in PM10 and PM2.5 aligns with the heating active hours,
suggesting that household coal consumption contributes to elevated PM
levels at both DZ and UB sites. In contrast, ZU and SS sites displayed
significantly lower annual PM10 and PM2.5 levels, with sudden frequent
spikes in spring followed by occasional instances in autumn. These
annual variations were more pronounced in PM10 compared to PM2.5,
highlighting the impact of Gobi dust and sand storms. Similar
occurrences were also noted at the DZ site, suggesting its exposure to
both winter heating emissions and natural spring dust, reflecting its
Gobi-region characteristics. The annual variability in PMs with higher
concentrations during nighttime and colder months, indicating the
influence of localized emission sources and reduced boundary layer
mixing at UB and DZ sites. Additionally, the upward extended ranges
without the bottom bottle of the violin plot, demonstrating greater
variability during colder months. It implies instability in
concentrations potentially reaching high levels above 400 \(\mu g/m^3\)
when Arctic oscillation/Siberian high intensifies with heating, and
dropping below 50 \(\mu g/m^3\) when it weakens or heating demand
decreases. These findings confirm the (DZ) emerging PM pattern is caused
by household activities and influenced by meteorological conditions.
Overall, meteorological factors appear to play a crucial role in
governing PM levels.

--\textgreater{} 3.2.1

\subsubsection{The emission patterns of interrelations among
meteorological variables at the study
sites}\label{the-emission-patterns-of-interrelations-among-meteorological-variables-at-the-study-sites}

\begin{itemize}
\tightlist
\item
  PMs varies with the wind speed and visibility
\item
  In general, three distinct patterns were resulted with PCA analysis,
  which is in consistent with temporal variation. explained with
  variables and changes in drivings (with PCA analysis)
\end{itemize}

{[}AND{]} To distinguish the emission driving variables, we demonstrated
interrelations between(WS) and visibility (VIS) among particulate
matters of PM10 and PM2.5 for study sites (figure\_4). {[}AND{]} Higher
PM concentrations tend to correspond to lower wind speeds (smaller
circles) and reduced visibility (darker blue points) in UB. {[}AND{]}
All Gobi sites show the largest variability in PM concentrations, with
extreme outliers correlated with the WS primarily in the springs.

{[}AND{]} In comparison between PM10 increase with the PM2.5, as
concentrations increase in PM10 concentration the ratio 1:1 is
inscreased approximately 1.3 times for UB, and 2 times for all three
Gobi sites. This increase in Gobi sites were further increased upto 5
times in ZU, and XXX times in SS sites showing the insufficient PM2.5
particulates in the those sites.

The Principal Component Analysis (PCA) biplot (Figure 7a) highlights the
relationships among key variables, including PM10, PM2.5, wind speed
(WS), wind direction (WD), visibility, relative humidity (r), and
optical properties of aerosols (OPC). The first two principal components
(Dim 1 and Dim 2) explain 66.52\% of the total variance, with Dim 1
(44.36\%) strongly associated with PM10 and PM2.5, indicating that these
pollutants are major contributors to air quality variations. Visibility
is inversely correlated with PM10 and PM2.5, suggesting that reduced
visibility aligns with higher pollution levels. Wind speed and wind
direction are positively aligned with Dim 2 (22.16\%), reflecting their
influence on pollutant dispersion.

The clustering of monitoring sites based on PM concentrations and
geographic characteristics (Figure 7b) reveals distinct patterns. UB
(urban background) is characterized by high PM concentrations (positive
Dim 1 scores), while SS (suburban site) demonstrates low PM levels,
clustering tightly along the negative Dim 1 axis. DZ (industrial site)
exhibits substantial variability, with clusters extending into higher
Dim 1 and Dim 2 ranges, reflecting the complex interplay of emission
sources and meteorological factors. ZU (rural site) overlaps with SS but
shows greater spread, indicating moderate pollution levels influenced by
seasonal and localized factors. These results underscore the significant
spatial and meteorological influences on air quality across the studied
locations.

{[}BUT{]}

{[}THEREFORE{]} 3 clusters, supply is different?

\subsubsection{The recent trends in concentrations of PMS and
fine-coarse fractional changes at the
sites}\label{the-recent-trends-in-concentrations-of-pms-and-fine-coarse-fractional-changes-at-the-sites}

Figure 8 illustrates the trend analysis of PM10 and PM2.5 concentrations
across four monitoring sites (UB, DZ, ZU, and SS) from 2009 to 2020. The
time series demonstrates seasonal patterns and trends, with notable
differences in particulate matter concentrations between the four
seasons: Q1 (winter), Q2 (spring), Q3 (summer), and Q4 (autumn).

The p-values displayed on the figure indicate the statistical
significance of trend changes in PM10 and PM2.5 levels for each season.
At UB, significant negative trends are observed for Q1 and Q2 (p
\textless{} 0.001). At DZ, a significant positive trend is observed for
Q1 (R = 0.17, p \textless{} 0.001), indicating increasing particulate
matter levels during this season.17, p \textless{} 0.001), while Q2
exhibits no significant trend (R = 0, p = 0.915)., highlighting a
decreasing pattern in particulate concentrations during these periods
over time. In contrast, Q3 and Q4 show weaker or nonsignificant trends
(e.g., R = -0.03, p \textgreater{} 0.2), suggesting minimal changes in
PM levels during the warmer months.

The industrial site (DZ) exhibits higher variability in PM levels, with
periodic spikes indicative of localized emissions and industrial
activities. Urban background (UB) consistently records elevated PM
levels during Q1, driven by heating and combustion-related sources.
Rural (ZU) and suburban (SS) sites generally display lower particulate
concentrations. However, at ZU, a significant negative trend is observed
in both PM10 and PM2.5 for Q1, suggesting a consistent reduction in
particulate matter during winter months (may reduced transboundary
traffic). At SS, a negative trend is observed in PM2.5 for Q2,
reflecting a seasonal decline potentially linked to reduced wind speed
during spring.

The significant decreasing trends in PM concentrations in UB during
colder months likely indicate improvements in emission control measures
and a transition to more efficient heating practices. In contrast, a
significant increasing trend is observed at DZ, particularly during Q1,
suggesting rising emissions potentially associated with urbanized
household practices and other local activities in Mongolia's regional
towns.

\begin{itemize}
\tightlist
\item
  There are significant changes in time-series of PMs at 4 seasons
\item
  There any significant changes in ratio in the spring in respect to
  winter in DZ.
\item
  Close relationships was found between PM2.5 in winter and r values in
  the spring.
\end{itemize}

\subsection{Conclusions}\label{conclusions}

In this study, we investigated the temporal variations of PM2.5 and PM10
concentrations at the 4 sites of rural and urban those located along the
the wind corridor. Three distinct variations has been detected.

Air quality in urban sites is episodically dictated by dust events in
spring or late autumn, yet seasonally governed by anthropogenic
emissions in winter. Air quality in rural sites of SS and ZU is
episodically dictated by dust events in spring or late autumn. Air
quality in rural sites of SS and ZU is episodically dictated by dust
events in spring or late autumn.

A clear seasonal variations in the sites of UB and DZ is {[}Air quality
is governed by natural dust emission, and anthropogenic emissions{]}

\begin{itemize}
\tightlist
\item
  Due to rapid increase in urban, and combustion of coal/oyutolgoi for
  heating winter conditions results a highly increase in not only
  capital city but also towns
\item
  In a result, spring coarse dust, plus winter fine pollutants {[}spring
  coarse dust is immediately transported and deposited in the source
  area, whereas winter fine pollutants is permanently stayed in the
  source area due to stagnant atmosphere govern over entire country.,
  perhaps float- ing in the near surface, deposits in the surface{]}
\item
  Alarms, the Mongolian dust in the spring, optical properties will be
  shifted; this gives \ldots{} Gobi dust and sand storms has become
  tuiren, from the shoroon shuurga. which clearly requires the
  attention.
\end{itemize}

Following problems

\begin{itemize}
\tightlist
\item
  On downwind regions
\item
  On national-level Demonstrating temporal and spatial variations of
\end{itemize}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_3.png}
\caption{Distinct concentrations of coarse and fine particulates among
sites}
\end{figure}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Compare the concentrations of PMs at UB is the 2. Significance level
  difference 3. Conclude
\end{enumerate}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.4375in,height=\textheight,keepaspectratio]{images/figure_4.png}
\caption{Annual variations of \$PM\_\{10\}\$ and \$PM\_\{2.5\}\$}
\end{figure}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Clear annual variations at UB and DZ from pm2.5 pollutions 2. at ZU,
  and SS has a seasonally peaks episodic spring and late autumn from
  PM10
\end{enumerate}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_5.png}
\caption{Daily variations of \(PM_{10}\) and \(PM_{2.5}\) at UB and DZ
sites}
\end{figure}

\newpage
\subsection{Meteorological influence on $PM_{10}$ and $PM_{2.5}$ variations}
\label{subsec2}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_6.png}
\caption{Relationships between meteorological major factors and
variations of \(PM_{10}\) and \(PM_{2.5}\)}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=\textheight,keepaspectratio]{images/figure_7.png}
\caption{Spatio-temporal distinct feature of variations of \(PM_{10}\)
and \(PM_{2.5}\) with PCA analysis}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.08333in,height=\textheight,keepaspectratio]{images/figure_7b.png}
\caption{Patterns of meteorology and PMs at the 4 sites}
\end{figure}

\newpage
\subsection{Trends}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_8.png}
\caption{Interannual and seasonal trends of \(PM_{10}\) and \(PM_{2.5}\)
variations}
\end{figure}

\newpage

\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_1.png}{]}

\begin{figure}
\centering
\includegraphics[width=4.16667in,height=\textheight,keepaspectratio]{images/table_1.png}
\caption{\textbf{Table 1}. A description of datasets obtained at the
sites}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/scheme_1.png}
\caption{Scheme 1. Data handling procedure}
\end{figure}

\newpage

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_2b.png}
\caption{Figure 2. Data gap filling}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{images/figure_2c.png}
\caption{Figure 2b. Data gap filling}
\end{figure}

(1)

\newpage

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-Munkh2017}
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Munkhtsetseg E}, \textbf{Shinoda M},
\textbf{Ishizuka M}, \textbf{Mikami M}, \textbf{Kimura R},
\textbf{Nikolich G}. 2017. Anthropogenic dust emissions due to livestock
trampling in a mongolian temperate grassland. Atmospheric Chemistry and
Physics \textbf{17}:11389--11401.
doi:\href{https://doi.org/10.5194/acp-17-11389-2017}{10.5194/acp-17-11389-2017}.}

\end{CSLReferences}

\end{document}

### 2024-11-22 submission/manuscript.Rmd
draft processing

@@
 
-Munkh\^1, Joeseph P. Schmo\^2, Sally J. Rivers\^1, Patrick D.
-Schloss^1^$\dagger$
+Munkhtsetseg\^1\^5$^*$, Shimizu. A\^2, Matsuki. A\^3, Batdorj. D\^4,
+Matsui. H^1$^*$
 
@@
 
-$\dagger$ To whom correspondence should be addressed:
-[pschloss\@umich.edu](mailto:pschloss@umich.edu){.email}
+$^*$ To whom correspondence should be addressed:
+[e.munkhtsetseg\@yahoo.com](mailto:e.munkhtsetseg@yahoo.com){.email}
 
-1\. Department of Microbiology and Immunology, University of Michigan,
-Ann Arbor, MI 48109
+1\. Nagoya University, Japan
 
-2\. Other department contact information
+2\. NIES
 
@@
 ### The spatio-temporal variations of the PMs at the study sites
-To comparatively examine the spatio-difference regarding to the emissions of PMs, we illustrated the hourly observed values of PM10 and PM2.5 for all sites (figure_3).
-On each site, PM10 is more sporadic than PM2.5 due to dust epidemic nature.
-In conjunction, the mean values averaged from hourly values are larger than its median concentrations both on PM10 and PM2.5 for all sites.
-The mean variables of p-value show that concentrations of particulate matters significantly differ with 99% confident levels at all sites (figure_3), except  95% confident level between DZ and UB on PM10 (figure_3a).
-It clearly exhibits general variations of particulate matters as expected and emphasizes that the significant difference in the concentration values among the sites.
-Besides all sites are quantitatively differ by the values of concentrations in particulate matters, there are two main characteristics can be seen when we compare the median deviations from its mean values etc.
+To evaluate the spatial variations in particulate matter (PM) concentrations, we displayed hourly observed values of PM10 and PM2.5 for all study sites (figure_3).
+The mean p-values indicate that PM concentrations differ significantly at a 99% confidence level across all sites (figure_3), with the exception of a 95% confidence level between DZ and UB for PM10 (figure_3a), highlighting substantial concentration disparities among sites.
+While quantitative differences in PM concentration values exist across all sites, two key patterns emerge when examining median deviations from mean values and irregular observation fluctuations.
+For instance, PM10 demonstrates more erratic behavior than PM2.5 at each location, particularly evident at ZU and SS sites.
+Furthermore, the mean values calculated from hourly measurements surpass the median concentrations for both PM10 and PM2.5 across all sites, with notable prominence at UB and DZ locations.
+Consequently, significant spatial differences in PM concentrations exist among all sites, regardless of urban or rural classification.
+However, the sites can be categorized into two groups based on their characteristics: UB (urban) and DZ (rural town, Gobi); and SS (rural, Gobi) and ZU (rural, Gobi).
+These findings for DZ appear to support our hypothesis of emerging new emission patterns related to increased coal consumption during winter months.
+---
+[AND] To examine the PM emerging patterns whether it related to household heating fuel, we demonstrated annual variations in PM10 and PM2.5 concentrations at the sites.
+[AND] Significant annual variations in PM10 and PM2.5 levels demonstrating higher concentrations in colder months and lower concentrations in warmer months at UB and DZ sites.
 
-- such patterns are strongly manifested in ZU and SS sites.  
-- the mean values those averaged from hourly concentrations of PM10 and PM2.5 is larger than its median are strongly manifested for all in UB and DZ sites.
-- Additionally, consider the significance of the ..
+[AND] During colder months (January, November, December), PM10 concentrations contributed by elevated levels in PM2.5 were exceed 100 $\mu g/m^3$, accompanied by PM10 and PM2.5 concentrations reached their lowest points during warmer months (May-September), with medians and ranges consistently below 50 $\mu g/m^3$.
+This distinct seasonal trend, peaking during the same cold months are supported by the diurnal variations in PM10 and PM2.5 concentrations at sites DZ and UB.
+A pronounced daily cycle is evident, at both DZ and UB sites, where PM concentrations peak during nighttime and early morning hours (approximately 8 PM to 4 AM UTC), with median values surpassing 50 $\mu g/m^3$.
+PM2.5 concentrations consistently follow a similar pattern but remain lower than PM10.
+In contrast, both pollutants exhibit reduced concentrations during daytime hours (8AM to 4PM UTC), likely due to increased atmospheric dispersion.
+At site UB, a comparable daily trend is observed, but with lower overall concentrations and less pronounced peaks.
+The variability, indicated by the interquartile range, is higher during nighttime, suggesting the influence of localized emission sources and reduced boundary layer mixing.
+These diurnal patterns underscore the temporal dynamics of air pollution, influenced by both anthropogenic activities and meteorological conditions.
 
-This highlight that the fine and coarse particulate matters significantly vary among sites of urban and rural sites, and even within Gobi (rural) sites.
-Moreover, .... has a diversely characteristically diversified in respect with the PM2.5 (fine) to PM10 (coarse) particulates.,. which requires urban impacts
+Analysis of PM10 and PM2.5 concentrations revealed that ZU and SS maintained significantly lower levels throughout the year, while UB and DZ exhibited higher concentrations, especially during colder months.
+The consistently lower particulate matter levels at ZU and SS suggest localized differences in pollution sources and dispersion patterns.
+These results underscore the seasonal and spatial variability in air quality, likely influenced by meteorological conditions and human activities.
+The observed seasonal patterns in PM10 and PM2.5 concentrations may be attributed to changes in weather conditions and anthropogenic activities.
+ --
+ high with violin meteorological effects
+--
+Annual maximum in the winter for DZ and UB are from PM2.5, which results an increase in PM10.
 
-To reveal the natural and anthropogenic impacts on the concentrations of particulate matters, we demonstrated annual variations of PM10 and PM2.5 for each sites (figure_4).
+clearly states that ....
+DZ and UB are from the ...
+household ., that concentration can be as high as a arctic oscillation/Siberian high intensifies with the heating, as low as it loose or may combination of heating demand drops.
+
+[BUT]
+
+[THEREFORE] PM10 and PM2.5 concentrations are larger in the spring followed by the autumn for SS and ZU sites.
+Annual PM10 and PM2.5 concentrations peaks in the winter aligned with the daily variations happens at the heating time for SS and ZU sites.
+This points that the increase in PM2.5 and PM10 is from the coal combustion.
+It requires the cause the behind such the variations. DZ site is polluted in the winter by the heating and in the spring by the natural dust.
+
 3.  Do PM2.5 particulates has contributed to the PM10 annual
@@
 
-[Therefore] SS and ZU sites are mainly affected by the spring dust, followed by the autumn dust. Annual maximum in the winter for DZ and UB are from PM2.5, which results an increase in PM10.
-It requires the cause the behind such the variations. DZ site is polluted in the winter by the heating and in the spring by the natural dust.
-
-[Therefore] Spatio-temporally in two class; consists of 2 Gobi sites, and 1 urban plus urbanized Gobi sites.
-
 ### The emission patterns of interrelations among meteorological variables at the study sites
 
-        -   PMs varies with the wind speed and visibility
-        -   In general, three distinct patterns were resulted with PCA
-            analysis, which is in consistent with temporal variation.
-            explained with variables and changes in drivings (with PCA
-            analysis)
+[AND] To distinguish the emission driving variables, at first we demonstrated interrelations between wind speed, visibility among particulate matters of PM10 and PM2.5 for each sites (figure_4).
+
+Figure 6 illustrates the relationship between PM10 and PM2.5 concentrations across the four sites (UB, DZ, ZU, and SS) for different seasons, alongside variations in wind speed (WS) and visibility (VIS).
+At all sites, a strong positive correlation between PM10 and PM2.5 is evident, as highlighted by the linear trendlines.
+The slopes suggest proportionality, with PM2.5 typically contributing a significant fraction of PM10 concentrations.
+
+Seasonal differences are prominent, with higher PM concentrations observed during November-February (marked as $"+"$ symbols), particularly at UB and DZ, where values significantly exceed those from other periods.
+Wind speed and visibility also exhibit notable interactions; higher PM concentrations tend to correspond to lower wind speeds (smaller circles) and reduced visibility (darker blue points).
+Conversely, during March–June and July–October, marked by circles and triangles, respectively, the overall concentrations are lower, especially at ZU and SS, indicating better air quality conditions likely due to more favorable meteorological factors.
+
+The site DZ shows the largest variability in PM concentrations, with extreme outliers during high-pollution periods.
+SS, on the other hand, exhibits relatively low PM levels across all seasons, aligning with its less polluted status.
+These results underscore the importance of local factors, including meteorological conditions and emission sources, in driving the observed PM dynamics.
 
 DZ site is polluted in the winter by the heating and in the spring by the natural dust.
+
+- PMs varies with the wind speed and visibility
+- In general, three distinct patterns were resulted with PCA analysis, which is in consistent with temporal variation. explained with variables and changes in drivings (with PCA analysis)
 

### 2024-11-22 submission/manuscript.tex
draft processing

@@
 \makeatother
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 
-Munkh\^{}1, Joeseph P. Schmo\^{}2, Sally J. Rivers\^{}1, Patrick D.
-Schloss\textsuperscript{1}\(\dagger\)
+Munkhtsetseg\^{}1\^{}5\(^*\), Shimizu. A\^{}2, Matsuki. A\^{}3, Batdorj.
+D\^{}4, Matsui. H\^{}1\(^*\)
 
@@
 
-\(\dagger\) To whom correspondence should be addressed:
-\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}
+\(^*\) To whom correspondence should be addressed:
+\href{mailto:e.munkhtsetseg@yahoo.com}{\nolinkurl{e.munkhtsetseg@yahoo.com}}
 
-1. Department of Microbiology and Immunology, University of Michigan,
-Ann Arbor, MI 48109
+1. Nagoya University, Japan
 
-2. Other department contact information
+2. NIES
 
@@
 
-\subsection{Abstract}\label{abstract}
+\subsection{Abstract (150 words)}\label{abstract-150-words}
 
@@
 \subsection{Introduction}\label{introduction}
+
+\begin{verbatim}
+* Advanced the knowledge of global dust, has reached to recognize the sources,.
+  - Classification dust brown color, seasonal characteristics, with coarse fractions.
+  - This knowledge further efficient to climate system when elaborating dust-aerosol effects.
+  - But, a large uncertainties in the global dust model has existed so for climate models which clearly limits our understanding the climate system and shape the facing global issues of global warming.
+  - This is mainly caused by the lack of parameterization and recognition of iterative changes controlled by the natural forces and anthropogenic drivings.
+
+* Mongolian dust brown color, seasonal characteristics, with coarse fractions.
+  - Mongolian dust has an attention of the its mass fraction in global dust, yet unlikely elaborated in the climate models due to its majority of coarse fraction for its a small contribution to the climate system through its radiative feedback.
+  - But, such recognized characterization might get no longer valid due to recent change in the driving of the emissions of air particulate matters.
+A large high concentrations of PM2.5 in the capital city of Mongolia has been observed  as a result of the heavy consumptions of coal as a winter heating has rapidly spread as a mining industry taken off since 2000.
+Winter weather stagnant conditions governed by the Siberian magnifies the concentrations of the particulate matter emissions by trapping the polluted air below the boundary layer, so that results in a very large high concentrations of PM2.5, locally.
+Even  recognized as one of the highly polluted capital cities in the world.
+  - Therefore, It is important to examine the emerging changes and shifting patterns of air particulate matters in Mongolia.
+More importantly, it is essential to reveal the significant changes in the the altered fraction particularly, in the dust seasons considering its high potential of intriuging in the free atmosphere to transported in the long-distance, so carrying capacity of the role to shift the global climate system, and its side impacts on downwind regions.
+\end{verbatim}
+
+\begin{itemize}
+\tightlist
+\item
+  Study goal
+
+  \begin{itemize}
+  \tightlist
+  \item
+    We hypothesize \ldots{}
+  \item
+    Our study will benefit not only to the global dust research but also
+    climate, and further to the country itself for urban planning, and
+    coal combustion.
+  \end{itemize}
+\end{itemize}
 
@@
 
-To comparatively examine the spatio-difference regarding to the
-emissions of PMs, we illustrated the hourly observed values of PM10 and
-PM2.5 for all sites (figure\_3). On each site, PM10 is more sporadic
-than PM2.5 due to dust epidemic nature. In conjunction, the mean values
-averaged from hourly values are larger than its median concentrations
-both on PM10 and PM2.5 for all sites. The mean variables of p-value show
-that concentrations of particulate matters significantly differ with
-99\% confident levels at all sites (figure\_3), except 95\% confident
-level between DZ and UB on PM10 (figure\_3a). It clearly exhibits
-general variations of particulate matters as expected and emphasizes
-that the significant difference in the concentration values among the
-sites. Besides all sites are quantitatively differ by the values of
-concentrations in particulate matters, there are two main
-characteristics can be seen when we compare the median deviations from
-its mean values etc.
+\subsection{To evaluate the spatial variations in particulate matter
+(PM) concentrations, we displayed hourly observed values of PM10 and
+PM2.5 for all study sites (figure\_3). The mean p-values indicate that
+PM concentrations differ significantly at a 99\% confidence level across
+all sites (figure\_3), with the exception of a 95\% confidence level
+between DZ and UB for PM10 (figure\_3a), highlighting substantial
+concentration disparities among sites. While quantitative differences in
+PM concentration values exist across all sites, two key patterns emerge
+when examining median deviations from mean values and irregular
+observation fluctuations. For instance, PM10 demonstrates more erratic
+behavior than PM2.5 at each location, particularly evident at ZU and SS
+sites. Furthermore, the mean values calculated from hourly measurements
+surpass the median concentrations for both PM10 and PM2.5 across all
+sites, with notable prominence at UB and DZ locations. Consequently,
+significant spatial differences in PM concentrations exist among all
+sites, regardless of urban or rural classification. However, the sites
+can be categorized into two groups based on their characteristics: UB
+(urban) and DZ (rural town, Gobi); and SS (rural, Gobi) and ZU (rural,
+Gobi). These findings for DZ appear to support our hypothesis of
+emerging new emission patterns related to increased coal consumption
+during winter
+months.}\label{to-evaluate-the-spatial-variations-in-particulate-matter-pm-concentrations-we-displayed-hourly-observed-values-of-pm10-and-pm2.5-for-all-study-sites-figure_3.-the-mean-p-values-indicate-that-pm-concentrations-differ-significantly-at-a-99-confidence-level-across-all-sites-figure_3-with-the-exception-of-a-95-confidence-level-between-dz-and-ub-for-pm10-figure_3a-highlighting-substantial-concentration-disparities-among-sites.-while-quantitative-differences-in-pm-concentration-values-exist-across-all-sites-two-key-patterns-emerge-when-examining-median-deviations-from-mean-values-and-irregular-observation-fluctuations.-for-instance-pm10-demonstrates-more-erratic-behavior-than-pm2.5-at-each-location-particularly-evident-at-zu-and-ss-sites.-furthermore-the-mean-values-calculated-from-hourly-measurements-surpass-the-median-concentrations-for-both-pm10-and-pm2.5-across-all-sites-with-notable-prominence-at-ub-and-dz-locations.-consequently-significant-spatial-differences-in-pm-concentrations-exist-among-all-sites-regardless-of-urban-or-rural-classification.-however-the-sites-can-be-categorized-into-two-groups-based-on-their-characteristics-ub-urban-and-dz-rural-town-gobi-and-ss-rural-gobi-and-zu-rural-gobi.-these-findings-for-dz-appear-to-support-our-hypothesis-of-emerging-new-emission-patterns-related-to-increased-coal-consumption-during-winter-months.}
 
-\begin{itemize}
+{[}AND{]} To examine the PM emerging patterns whether it related to
+household heating fuel, we demonstrated annual variations in PM10 and
+PM2.5 concentrations at the sites. {[}AND{]} Significant annual
+variations in PM10 and PM2.5 levels demonstrating higher concentrations
+in colder months and lower concentrations in warmer months at UB and DZ
+sites.
+
+{[}AND{]} During colder months (January, November, December), PM10
+concentrations contributed by elevated levels in PM2.5 were exceed 100
+\(\mu g/m^3\), accompanied by PM10 and PM2.5 concentrations reached
+their lowest points during warmer months (May-September), with medians
+and ranges consistently below 50 \(\mu g/m^3\). This distinct seasonal
+trend, peaking during the same cold months are supported by the diurnal
+variations in PM10 and PM2.5 concentrations at sites DZ and UB. A
+pronounced daily cycle is evident, at both DZ and UB sites, where PM
+concentrations peak during nighttime and early morning hours
+(approximately 8 PM to 4 AM UTC), with median values surpassing 50
+\(\mu g/m^3\). PM2.5 concentrations consistently follow a similar
+pattern but remain lower than PM10. In contrast, both pollutants exhibit
+reduced concentrations during daytime hours (8AM to 4PM UTC), likely due
+to increased atmospheric dispersion. At site UB, a comparable daily
+trend is observed, but with lower overall concentrations and less
+pronounced peaks. The variability, indicated by the interquartile range,
+is higher during nighttime, suggesting the influence of localized
+emission sources and reduced boundary layer mixing. These diurnal
+patterns underscore the temporal dynamics of air pollution, influenced
+by both anthropogenic activities and meteorological conditions.
+
+\begin{longtable}[]{@{}l@{}}
+\toprule\noalign{}
+nalysis of PM10 and PM2.5 concentrations revealed that ZU and SS
+maintained significantly lower levels throughout the year, while UB and
+DZ exhibited higher concentrations, especially during colder months. The
+consistently lower particulate matter levels at ZU and SS suggest
+localized differences in pollution sources and dispersion patterns.
+These results underscore the seasonal and spatial variability in air
+quality, likely influenced by meteorological conditions and human
+activities. The observed seasonal patterns in PM10 and PM2.5
+concentrations may be attributed to changes in weather conditions and
+anthropogenic activities. \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+high with violin meteorological effects \\
+\end{longtable}
+
+Annual maximum in the winter for DZ and UB are from PM2.5, which results
+an increase in PM10.
+
+clearly states that \ldots. DZ and UB are from the \ldots{} household .,
+that concentration can be as high as a arctic oscillation/Siberian high
+intensifies with the heating, as low as it loose or may combination of
+heating demand drops.
+
+{[}BUT{]}
+
+{[}THEREFORE{]} PM10 and PM2.5 concentrations are larger in the spring
+followed by the autumn for SS and ZU sites. Annual PM10 and PM2.5
+concentrations peaks in the winter aligned with the daily variations
+happens at the heating time for SS and ZU sites. This points that the
+increase in PM2.5 and PM10 is from the coal combustion. It requires the
+cause the behind such the variations. DZ site is polluted in the winter
+by the heating and in the spring by the natural dust.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\setcounter{enumi}{2}
 \tightlist
 \item
-  such patterns are strongly manifested in ZU and SS sites.\\
-\item
-  the mean values those averaged from hourly concentrations of PM10 and
-  PM2.5 is larger than its median are strongly manifested for all in UB
-  and DZ sites.
-\item
-  Additionally, consider the significance of the ..
-\end{itemize}
-
-This highlight that the fine and coarse particulate matters
-significantly vary among sites of urban and rural sites, and even within
-Gobi (rural) sites. Moreover, \ldots. has a diversely characteristically
-diversified in respect with the PM2.5 (fine) to PM10 (coarse)
-particulates.,. which requires urban impacts
-
-To reveal the natural and anthropogenic impacts on the concentrations of
-particulate matters, we demonstrated annual variations of PM10 and PM2.5
-for each sites (figure\_4). 3. Do PM2.5 particulates has contributed to
-the PM10 annual variations? Distinct temporal variations has existed
-among the sites. - PM2.5 particulates has contributed to the PM10 annual
-variations in UB and in DZ in winter. - If yes, how much, and when and
-where? - What is the sd, mean, and median - box plot - violin - scatter
-points, epidemic, sporadic - Daily variations to examine it related to
-the heating - 2 peaks: smaller and bigger - compare the t-duration
-exceeds 50mug/m3/hour
-
-{[}Therefore{]} SS and ZU sites are mainly affected by the spring dust,
-followed by the autumn dust. Annual maximum in the winter for DZ and UB
-are from PM2.5, which results an increase in PM10. It requires the cause
-the behind such the variations. DZ site is polluted in the winter by the
-heating and in the spring by the natural dust.
-
-{[}Therefore{]} Spatio-temporally in two class; consists of 2 Gobi
-sites, and 1 urban plus urbanized Gobi sites.
+  Do PM2.5 particulates has contributed to the PM10 annual variations?
+  Distinct temporal variations has existed among the sites. - PM2.5
+  particulates has contributed to the PM10 annual variations in UB and
+  in DZ in winter. - If yes, how much, and when and where? - What is the
+  sd, mean, and median - box plot - violin - scatter points, epidemic,
+  sporadic - Daily variations to examine it related to the heating - 2
+  peaks: smaller and bigger - compare the t-duration exceeds
+  50mug/m3/hour
+\end{enumerate}
 
@@
 
-\begin{verbatim}
-    -   PMs varies with the wind speed and visibility
-    -   In general, three distinct patterns were resulted with PCA
-        analysis, which is in consistent with temporal variation.
-        explained with variables and changes in drivings (with PCA
-        analysis)
-\end{verbatim}
+{[}AND{]} To distinguish the emission driving variables, at first we
+demonstrated interrelations between wind speed, visibility among
+particulate matters of PM10 and PM2.5 for each sites (figure\_4).
+
+Figure 6 illustrates the relationship between PM10 and PM2.5
+concentrations across the four sites (UB, DZ, ZU, and SS) for different
+seasons, alongside variations in wind speed (WS) and visibility (VIS).
+At all sites, a strong positive correlation between PM10 and PM2.5 is
+evident, as highlighted by the linear trendlines. The slopes suggest
+proportionality, with PM2.5 typically contributing a significant
+fraction of PM10 concentrations.
+
+Seasonal differences are prominent, with higher PM concentrations
+observed during November-February (marked as \("+"\) symbols),
+particularly at UB and DZ, where values significantly exceed those from
+other periods. Wind speed and visibility also exhibit notable
+interactions; higher PM concentrations tend to correspond to lower wind
+speeds (smaller circles) and reduced visibility (darker blue points).
+Conversely, during March--June and July--October, marked by circles and
+triangles, respectively, the overall concentrations are lower,
+especially at ZU and SS, indicating better air quality conditions likely
+due to more favorable meteorological factors.
+
+The site DZ shows the largest variability in PM concentrations, with
+extreme outliers during high-pollution periods. SS, on the other hand,
+exhibits relatively low PM levels across all seasons, aligning with its
+less polluted status. These results underscore the importance of local
+factors, including meteorological conditions and emission sources, in
+driving the observed PM dynamics.
 
@@
 the natural dust.
+
+\begin{itemize}
+\tightlist
+\item
+  PMs varies with the wind speed and visibility
+\item
+  In general, three distinct patterns were resulted with PCA analysis,
+  which is in consistent with temporal variation. explained with
+  variables and changes in drivings (with PCA analysis)
+\end{itemize}
 
@@
 
+\subsubsection{Materials}\label{materials}
+
+\subsubsection{Methods 3,000 words}\label{methods-3000-words}
+
+\subsection{Acknowledgements}\label{acknowledgements}
+
+Keep acknowledgements brief and do not include thanks to anonymous
+referees or editors, or effusive comments. Grant or contribution numbers
+may be acknowledged.
+
+\subsection{Figures (10)}\label{figures-10}
+
+Figure legends should be \textless350 words each. They should begin with
+a brief title sentence for the whole figure and continue with a short
+statement of what is depicted in the figure, not the results (or data)
+of the experiment or the methods used. Legends should be detailed enough
+so that each figure and caption can, as far as possible, be understood
+in isolation from the main text.
+
+Tables. Each table should be prepared using the Table menu in Word or
+the table environment in TeX/LaTeX and accompanied by a short title
+sentence describing what the table shows. Further details can be
+included as footnotes to the table.
+
+\subsection{References (70)}\label{references-70}
+
+\subsection{Supplementary}\label{supplementary}
+
+Author contributions. You must include a statement that specifies the
+individual contributions of each co-author. For example: ``A.P.M.
+`contributed' Y and Z; B.T.R. `contributed' Y,'' etc. See our authorship
+policies for more details.
+
+Competing interests. Submission of a competing interests statement is
+required for all content of the journal.
+
+Materials \& Correspondence. Indicate the author(s) to whom
+correspondence and material requests should be addressed.
+
+Supplementary information Please submit supplementary figures, small
+tables and text as a single combined PDF document. Tables longer than
+one page should be provided as an Excel or similar file type. For
+optimal quality video files please use H.264 encoding, the standard
+aspect ratio of 16:9 (4:3 is second best) and do not compress the video.
+We encourage submission of step-by-step synthesis procedures for
+chemical compounds and data on compound characterization. Supplementary
+information is not copyedited, so please ensure that it is clearly and
+succinctly presented, and that the style and terminology conform to the
+rest of the manuscript.
+
+\subsection{Materials and Methods}\label{materials-and-methods-1}
+
 \subsubsection{A description of study

### 2024-11-25 submission/Outline_20241121.Rmd
outline - title changed, abstract started, results drafted except the last figures

@@
 ---
-title: "**Spatio-temporal distinct patterns in variations of $PM_{10}$ and $PM_{2.5}$
-  relative to the recent drivings of emission sources in Mongolia**"
+title: "**Evidence for a Change to Aerosol Fractions: In a Gobi Town**"
 bibliography: references.bib
@@
 ## Abstract (150 words)
-
+Driven by mining industrial progress, household coal combustion has increased greatly in Mongolia since the 2000s.
+Demographic evidence has revealed an ongoing reduction in rural-nomadic lifestyle and a rise of the population at urban areas, which may gradually extend a household major source of fine particulate matter (PM2.5) air pollution in winter as a heating.
+Together with observations that lifespan in various animal species is flexible and can be increased by genetic or pharmaceutical intervention, these results have led to suggestions that longevity may not be subject to strict, species-specific genetic constraints.
+Here, by analysing global demographic data, we show that improvements in survival with age tend to decline after age 100, and that the age at death of the world’s oldest person has not increased since the 1990s.
+Our results strongly suggest that the maximum lifespan of humans is fixed and subject to natural constraints.
 Storyline:
@@
     fine particulates as well.
-
+    Spatio-temporal distinct patterns in variations of $PM_{10}$ and $PM_{2.5}$
+      relative to the recent drivings of emission sources in Mongolia
 \newpage
@@
 
---> 3.1.2
-
   To investigate whether emerging PM patterns are associated with household winter heating activities, we demonstrated annual variations in PM10 and PM2.5 concentrations at the sites.
@@
 
---> 3.2.1
+### The emission patterns of interrelations among meteorological variables at the study sites
+add table, add r 1.3, 2 on figure_6, add supplement figures
 
-### The emission patterns of interrelations among meteorological variables at the study sites
-  - PMs varies with the wind speed and visibility
-  - In general, three distinct patterns were resulted with PCA analysis, which is in consistent with temporal variation. explained with variables and changes in drivings (with PCA analysis)
+  To identify the key factors influencing PM emissions, we examined the relationships between wind speed (WS), visibility (VIS), and particulate matter (PM10 and PM2.5) concentrations across the study sites (figure_6).
+In UB during winter, elevated PM levels typically coincide with low wind speeds and reduced visibility (indicated by darker blue points).
+Notably, at DZ and ZU locations, high PM10 concentrations were observed during both low and high wind speed conditions in winter.
+All Gobi sites exhibit the greatest variation in PM concentrations during spring, with extreme outliers primarily associated with increased WS and reduced visibility below 10000 km.
+Similar findings were also revealed through Principal Component Analysis (PCA).
+The initial two principal components (Dim 1 and Dim 2) account for 66.52% of the total variance.
+Dim 1 shows a strong association with PM10 and PM2.5, while visibility demonstrates an inverse correlation, suggesting that reduced visibility corresponds to higher pollution levels.
+Wind speed and direction align positively with Dim 2 (22.16%), reflecting their impact on emissions.
+When comparing PM10 aligns with PM2.5, the 1:1 ratio increases approximately 1.3 times for UB, 2 times for DZ, XXX for ZU, and XXX for SS as PM10 concentrations increase.
+This increase in Gobi sites further escalated up to 2-5 times in ZU, and XXX times in SS sites, indicating insufficient PM2.5 particulates in those locations, particularly at the ZU site.
+It is worth noting that instances where PM2.5 values exceed PM10 reflect equipment accuracy, correction errors, and higher sensitivity rates (ability to record emissions such as smoking near the sensor area).
+However, this discrepancy diminishes as PM concentrations increase, validating that higher observational records of PM2.5 compared to PM10 do not invalidate our results.
+Furthermore, clustering of monitoring sites based on PM concentrations and geographic features (Figure 7b) reveals distinct patterns.
+UB (urban, capital city) is characterized by high PM concentrations (positive Dim 1 scores), while SS (rural, Gobi, town (site located in the prevailing wind above the town)) shows low PM levels, clustering tightly along the negative Dim 1 axis.
+DZ (rural, Gobi, town center) displays considerable variability, with clusters extending into higher Dim 1 and Dim 2 ranges, reflecting the complex interplay of emission sources and meteorological factors.
+ZU (rural, Gobi, village) overlaps with SS but exhibits greater spread, indicating moderate pollution levels influenced by seasonal and localized factors.
+These findings highlight the complex interplay of spatial, meteorological, and local factors influencing air particulate matter concentrations across the studied locations and emphasizing the need for considering not only regional influences but also site-specific characteristics.
 
-[AND] To distinguish the emission driving variables, we demonstrated interrelations between(WS) and visibility (VIS) among particulate matters of PM10 and PM2.5 for study sites (figure_4).
-[AND] Higher PM concentrations tend to correspond to lower wind speeds (smaller circles) and reduced visibility (darker blue points) in UB.
-[AND] All Gobi sites show the largest variability in PM concentrations, with extreme outliers correlated with the WS primarily in the springs.
-
-[AND] In comparison between PM10 increase with the PM2.5, as concentrations increase in PM10 concentration the ratio 1:1 is inscreased approximately 1.3 times for UB, and 2 times for all three Gobi sites.
-This increase in Gobi sites were further increased upto 5 times in ZU, and XXX times in SS sites showing the insufficient PM2.5 particulates in the those sites.
-
-The Principal Component Analysis (PCA) biplot (Figure 7a) highlights the relationships among key variables, including PM10, PM2.5, wind speed (WS), wind direction (WD), visibility, relative humidity (r), and optical properties of aerosols (OPC).
-The first two principal components (Dim 1 and Dim 2) explain 66.52% of the total variance, with Dim 1 (44.36%) strongly associated with PM10 and PM2.5, indicating that these pollutants are major contributors to air quality variations.
-Visibility is inversely correlated with PM10 and PM2.5, suggesting that reduced visibility aligns with higher pollution levels.
+fig caption:  the Principal Component Analysis (PCA) bi-plot (Figure 7a) highlights the relationships among key variables, including PM10, PM2.5, wind speed (WS), wind direction (WD), visibility, ratio of PM2.5 to PM10 (r), and numbers of aerosols by optical particle counter (OPC).
+The first two principal components (Dim 1 and Dim 2) explain 66.52% of the total variance, with Dim 1 (44.36%) strongly associated with PM10 and PM2.5.
 Wind speed and wind direction are positively aligned with Dim 2 (22.16%), reflecting their influence on pollutant dispersion.
 
-The clustering of monitoring sites based on PM concentrations and geographic characteristics (Figure 7b) reveals distinct patterns.
-UB (urban background) is characterized by high PM concentrations (positive Dim 1 scores), while SS (suburban site) demonstrates low PM levels, clustering tightly along the negative Dim 1 axis.
-DZ (industrial site) exhibits substantial variability, with clusters extending into higher Dim 1 and Dim 2 ranges, reflecting the complex interplay of emission sources and meteorological factors.
-ZU (rural site) overlaps with SS but shows greater spread, indicating moderate pollution levels influenced by seasonal and localized factors.
-These results underscore the significant spatial and meteorological influences on air quality across the studied locations.
+###  The recent trends in concentrations of PMS and fine-coarse fractional changes at the sites
+add table, add trend figure: r, add relationship figure: r in spring and pm2.5 in winter
 
-[BUT]
-
-[THEREFORE] 3 clusters, supply is different?
-
-###  The recent trends in concentrations of PMS and fine-coarse fractional changes at the sites
-Figure 8 illustrates the trend analysis of PM10 and PM2.5 concentrations across four monitoring sites (UB, DZ, ZU, and SS) from 2009 to 2020.
-The time series demonstrates seasonal patterns and trends, with notable differences in particulate matter concentrations between the four seasons: Q1 (winter), Q2 (spring), Q3 (summer), and Q4 (autumn).
-
-The p-values displayed on the figure indicate the statistical significance of trend changes in PM10 and PM2.5 levels for each season.
-At UB, significant negative trends are observed for Q1 and Q2 (p < 0.001).
-At DZ, a significant positive trend is observed for Q1 (R = 0.17, p < 0.001), indicating increasing particulate matter levels during this season.17, p < 0.001), while Q2 exhibits no significant trend (R = 0, p = 0.915)., highlighting a decreasing pattern in particulate concentrations during these periods over time.
-In contrast, Q3 and Q4 show weaker or nonsignificant trends (e.g., R = -0.03, p > 0.2), suggesting minimal changes in PM levels during the warmer months.
-
-The industrial site (DZ) exhibits higher variability in PM levels, with periodic spikes indicative of localized emissions and industrial activities.
-Urban background (UB) consistently records elevated PM levels during Q1, driven by heating and combustion-related sources.
-Rural (ZU) and suburban (SS) sites generally display lower particulate concentrations.
-However, at ZU, a significant negative trend is observed in both PM10 and PM2.5 for Q1, suggesting a consistent reduction in particulate matter during winter months (may reduced transboundary traffic).
-At SS, a negative trend is observed in PM2.5 for Q2, reflecting a seasonal decline potentially linked to  reduced wind speed during spring.
-
+  Figure 8 illustrates the trend analysis of PM10 and PM2.5 concentrations across study sites from 2009 to 2020.
+The time series demonstrates seasonal patterns and trends, with p-values of trend changes, seasonally: winter (Q1), spring (Q2), and summer-autumn (Q3).
+At UB, significant negative trends both in PM10 and PM2.5 concentrations are observed for winter and spring (Q1 and Q2, p < 0.001).
+At DZ, a significant positive trend in PM10 and PM2.5 concentrations is observed for winter (Q1, p < 0.001), indicating increasing particulate matter levels during this season.
+At ZU, a significant negative trend is observed in both PM10 and PM2.5 for winter (Q1, p < 0.001), suggesting a consistent reduction in particulate matter during winter months (may reduced transboundary traffic associated with covid periods, resulted in declined PM10 concentrations; or data gaps).
+At SS, a negative trend is observed in PM2.5 for spring (Q2, p < 0.001), reflecting a seasonal decline potentially linked to  reduced wind speed during spring.
 The significant decreasing trends in PM concentrations in UB during colder months likely indicate improvements in emission control measures and a transition to more efficient heating practices.
-In contrast, a significant increasing trend is observed at DZ, particularly during Q1, suggesting rising emissions potentially associated with urbanized  household practices and other local activities in Mongolia's regional towns.
-
-  - There are significant changes in time-series of PMs at 4 seasons
-  - There any significant changes in ratio in the spring in respect to winter in DZ.
-  - Close relationships was found between PM2.5 in winter and r values in the spring.
+In contrast, a significant increasing trend is observed at DZ, particularly during winter (Q1), suggesting rising emissions potentially associated with urbanized  household practices and other local activities in Mongolia's regional towns.
 
 ## Conclusions
+In this study, we investigated the temporal variations of PM2.5 and PM10 concentrations at the 4 sites of rural and urban those located along the
+the wind corridor.
+Air particulate matter concentrations in urban-town sites of UB and DZ is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
+Air particulate matter concentrations in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
 
-In this study, we investigated the temporal variations of PM2.5 and PM10
-concentrations at the 4 sites of rural and urban those located along the
-the wind corridor. Three distinct variations has been detected.
+Three distinct variations has been detected.
 
-Air quality in urban sites is episodically dictated by dust events in spring or late autumn, yet seasonally governed by anthropogenic emissions in winter.
-Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
-Air quality in rural sites of SS and ZU is episodically dictated by dust events in spring or late autumn.
-
-A clear seasonal variations in the sites of UB and DZ is [Air quality is
-governed by natural dust emission, and anthropogenic emissions]
+1. It is evident of the new emission patterns in Mongolia.
+2. Related to the winter emission patterns, fine particulates fraction in the spring is increased.
+3. This alarms...
+4.
+National level; meteorological impact is large.
+...
+However, reduced...
+On the other hand, it is ...
+with the towns.
+This points that air quality will be poor whether it is changed fuel, ..
+unless to change heating system.
+Therefore, it is not the reason to move the capital city.
+Only solution is to change the heating system, do not burn any type of the coal.
 

### 2024-11-25 submission/Outline_20241121.tex
outline - title changed, abstract started, results drafted except the last figures

@@
 \hypersetup{
-  pdftitle={Spatio-temporal distinct patterns in variations of PM\_\{10\} and PM\_\{2.5\} relative to the recent drivings of emission sources in Mongolia},
+  pdftitle={Evidence for a Change to Aerosol Fractions: In a Gobi Town},
   hidelinks,
@@
 
-\title{\textbf{Spatio-temporal distinct patterns in variations of
-\(PM_{10}\) and \(PM_{2.5}\) relative to the recent drivings of emission
-sources in Mongolia}}
+\title{\textbf{Evidence for a Change to Aerosol Fractions: In a Gobi
+Town}}
 \author{}
@@
 
-Storyline:
+Driven by mining industrial progress, household coal combustion has
+increased greatly in Mongolia since the 2000s. Demographic evidence has
+revealed an ongoing reduction in rural-nomadic lifestyle and a rise of
+the population at urban areas, which may gradually extend a household
+major source of fine particulate matter (PM2.5) air pollution in winter
+as a heating. Together with observations that lifespan in various animal
+species is flexible and can be increased by genetic or pharmaceutical
+intervention, these results have led to suggestions that longevity may
+not be subject to strict, species-specific genetic constraints. Here, by
+analysing global demographic data, we show that improvements in survival
+with age tend to decline after age 100, and that the age at death of the
+world's oldest person has not increased since the 1990s. Our results
+strongly suggest that the maximum lifespan of humans is fixed and
+subject to natural constraints. Storyline:
 
@@
   r ratio shows \ldots{} emission source; dust might carry anthropogenic
-  fine particulates as well.
+  fine particulates as well. Spatio-temporal distinct patterns in
+  variations of \(PM_{10}\) and \(PM_{2.5}\) relative to the recent
+  drivings of emission sources in Mongolia \newpage
 \end{enumerate}
-
-\newpage
 
@@
 
---\textgreater{} 3.1.2
-
 To investigate whether emerging PM patterns are associated with
@@
 
---\textgreater{} 3.2.1
-
 \subsubsection{The emission patterns of interrelations among
@@
 
-\begin{itemize}
-\tightlist
-\item
-  PMs varies with the wind speed and visibility
-\item
-  In general, three distinct patterns were resulted with PCA analysis,
-  which is in consistent with temporal variation. explained with
-  variables and changes in drivings (with PCA analysis)
-\end{itemize}
+add table, add r 1.3, 2 on figure\_6, add supplement figures
 
-{[}AND{]} To distinguish the emission driving variables, we demonstrated
-interrelations between(WS) and visibility (VIS) among particulate
-matters of PM10 and PM2.5 for study sites (figure\_4). {[}AND{]} Higher
-PM concentrations tend to correspond to lower wind speeds (smaller
-circles) and reduced visibility (darker blue points) in UB. {[}AND{]}
-All Gobi sites show the largest variability in PM concentrations, with
-extreme outliers correlated with the WS primarily in the springs.
+To identify the key factors influencing PM emissions, we examined the
+relationships between wind speed (WS), visibility (VIS), and particulate
+matter (PM10 and PM2.5) concentrations across the study sites
+(figure\_6). In UB during winter, elevated PM levels typically coincide
+with low wind speeds and reduced visibility (indicated by darker blue
+points). Notably, at DZ and ZU locations, high PM10 concentrations were
+observed during both low and high wind speed conditions in winter. All
+Gobi sites exhibit the greatest variation in PM concentrations during
+spring, with extreme outliers primarily associated with increased WS and
+reduced visibility below 10000 km. Similar findings were also revealed
+through Principal Component Analysis (PCA). The initial two principal
+components (Dim 1 and Dim 2) account for 66.52\% of the total variance.
+Dim 1 shows a strong association with PM10 and PM2.5, while visibility
+demonstrates an inverse correlation, suggesting that reduced visibility
+corresponds to higher pollution levels. Wind speed and direction align
+positively with Dim 2 (22.16\%), reflecting their impact on emissions.
+When comparing PM10 aligns with PM2.5, the 1:1 ratio increases
+approximately 1.3 times for UB, 2 times for DZ, XXX for ZU, and XXX for
+SS as PM10 concentrations increase. This increase in Gobi sites further
+escalated up to 2-5 times in ZU, and XXX times in SS sites, indicating
+insufficient PM2.5 particulates in those locations, particularly at the
+ZU site. It is worth noting that instances where PM2.5 values exceed
+PM10 reflect equipment accuracy, correction errors, and higher
+sensitivity rates (ability to record emissions such as smoking near the
+sensor area). However, this discrepancy diminishes as PM concentrations
+increase, validating that higher observational records of PM2.5 compared
+to PM10 do not invalidate our results. Furthermore, clustering of
+monitoring sites based on PM concentrations and geographic features
+(Figure 7b) reveals distinct patterns. UB (urban, capital city) is
+characterized by high PM concentrations (positive Dim 1 scores), while
+SS (rural, Gobi, town (site located in the prevailing wind above the
+town)) shows low PM levels, clustering tightly along the negative Dim 1
+axis. DZ (rural, Gobi, town center) displays considerable variability,
+with clusters extending into higher Dim 1 and Dim 2 ranges, reflecting
+the complex interplay of emission sources and meteorological factors. ZU
+(rural, Gobi, village) overlaps with SS but exhibits greater spread,
+indicating moderate pollution levels influenced by seasonal and
+localized factors. These findings highlight the complex interplay of
+spatial, meteorological, and local factors influencing air particulate
+matter concentrations across the studied locations and emphasizing the
+need for considering not only regional influences but also site-specific
+characteristics.
 
-{[}AND{]} In comparison between PM10 increase with the PM2.5, as
-concentrations increase in PM10 concentration the ratio 1:1 is
-inscreased approximately 1.3 times for UB, and 2 times for all three
-Gobi sites. This increase in Gobi sites were further increased upto 5
-times in ZU, and XXX times in SS sites showing the insufficient PM2.5
-particulates in the those sites.
-
-The Principal Component Analysis (PCA) biplot (Figure 7a) highlights the
-relationships among key variables, including PM10, PM2.5, wind speed
-(WS), wind direction (WD), visibility, relative humidity (r), and
-optical properties of aerosols (OPC). The first two principal components
-(Dim 1 and Dim 2) explain 66.52\% of the total variance, with Dim 1
-(44.36\%) strongly associated with PM10 and PM2.5, indicating that these
-pollutants are major contributors to air quality variations. Visibility
-is inversely correlated with PM10 and PM2.5, suggesting that reduced
-visibility aligns with higher pollution levels. Wind speed and wind
-direction are positively aligned with Dim 2 (22.16\%), reflecting their
-influence on pollutant dispersion.
-
-The clustering of monitoring sites based on PM concentrations and
-geographic characteristics (Figure 7b) reveals distinct patterns. UB
-(urban background) is characterized by high PM concentrations (positive
-Dim 1 scores), while SS (suburban site) demonstrates low PM levels,
-clustering tightly along the negative Dim 1 axis. DZ (industrial site)
-exhibits substantial variability, with clusters extending into higher
-Dim 1 and Dim 2 ranges, reflecting the complex interplay of emission
-sources and meteorological factors. ZU (rural site) overlaps with SS but
-shows greater spread, indicating moderate pollution levels influenced by
-seasonal and localized factors. These results underscore the significant
-spatial and meteorological influences on air quality across the studied
-locations.
-
-{[}BUT{]}
-
-{[}THEREFORE{]} 3 clusters, supply is different?
+fig caption: the Principal Component Analysis (PCA) bi-plot (Figure 7a)
+highlights the relationships among key variables, including PM10, PM2.5,
+wind speed (WS), wind direction (WD), visibility, ratio of PM2.5 to PM10
+(r), and numbers of aerosols by optical particle counter (OPC). The
+first two principal components (Dim 1 and Dim 2) explain 66.52\% of the
+total variance, with Dim 1 (44.36\%) strongly associated with PM10 and
+PM2.5. Wind speed and wind direction are positively aligned with Dim 2
+(22.16\%), reflecting their influence on pollutant dispersion.
 
@@
 
+add table, add trend figure: r, add relationship figure: r in spring and
+pm2.5 in winter
+
 Figure 8 illustrates the trend analysis of PM10 and PM2.5 concentrations
-across four monitoring sites (UB, DZ, ZU, and SS) from 2009 to 2020. The
-time series demonstrates seasonal patterns and trends, with notable
-differences in particulate matter concentrations between the four
-seasons: Q1 (winter), Q2 (spring), Q3 (summer), and Q4 (autumn).
-
-The p-values displayed on the figure indicate the statistical
-significance of trend changes in PM10 and PM2.5 levels for each season.
-At UB, significant negative trends are observed for Q1 and Q2 (p
-\textless{} 0.001). At DZ, a significant positive trend is observed for
-Q1 (R = 0.17, p \textless{} 0.001), indicating increasing particulate
-matter levels during this season.17, p \textless{} 0.001), while Q2
-exhibits no significant trend (R = 0, p = 0.915)., highlighting a
-decreasing pattern in particulate concentrations during these periods
-over time. In contrast, Q3 and Q4 show weaker or nonsignificant trends
-(e.g., R = -0.03, p \textgreater{} 0.2), suggesting minimal changes in
-PM levels during the warmer months.
-
-The industrial site (DZ) exhibits higher variability in PM levels, with
-periodic spikes indicative of localized emissions and industrial
-activities. Urban background (UB) consistently records elevated PM
-levels during Q1, driven by heating and combustion-related sources.
-Rural (ZU) and suburban (SS) sites generally display lower particulate
-concentrations. However, at ZU, a significant negative trend is observed
-in both PM10 and PM2.5 for Q1, suggesting a consistent reduction in
-particulate matter during winter months (may reduced transboundary
-traffic). At SS, a negative trend is observed in PM2.5 for Q2,
-reflecting a seasonal decline potentially linked to reduced wind speed
-during spring.
-
-The significant decreasing trends in PM concentrations in UB during
-colder months likely indicate improvements in emission control measures
-and a transition to more efficient heating practices. In contrast, a
-significant increasing trend is observed at DZ, particularly during Q1,
-suggesting rising emissions potentially associated with urbanized
-household practices and other local activities in Mongolia's regional
-towns.
-
-\begin{itemize}
-\tightlist
-\item
-  There are significant changes in time-series of PMs at 4 seasons
-\item
-  There any significant changes in ratio in the spring in respect to
-  winter in DZ.
-\item
-  Close relationships was found between PM2.5 in winter and r values in
-  the spring.
-\end{itemize}
+across study sites from 2009 to 2020. The time series demonstrates
+seasonal patterns and trends, with p-values of trend changes,
+seasonally: winter (Q1), spring (Q2), and summer-autumn (Q3). At UB,
+significant negative trends both in PM10 and PM2.5 concentrations are
+observed for winter and spring (Q1 and Q2, p \textless{} 0.001). At DZ,
+a significant positive trend in PM10 and PM2.5 concentrations is
+observed for winter (Q1, p \textless{} 0.001), indicating increasing
+particulate matter levels during this season. At ZU, a significant
+negative trend is observed in both PM10 and PM2.5 for winter (Q1, p
+\textless{} 0.001), suggesting a consistent reduction in particulate
+matter during winter months (may reduced transboundary traffic
+associated with covid periods, resulted in declined PM10 concentrations;
+or data gaps). At SS, a negative trend is observed in PM2.5 for spring
+(Q2, p \textless{} 0.001), reflecting a seasonal decline potentially
+linked to reduced wind speed during spring. The significant decreasing
+trends in PM concentrations in UB during colder months likely indicate
+improvements in emission control measures and a transition to more
+efficient heating practices. In contrast, a significant increasing trend
+is observed at DZ, particularly during winter (Q1), suggesting rising
+emissions potentially associated with urbanized household practices and
+other local activities in Mongolia's regional towns.
 
@@
 concentrations at the 4 sites of rural and urban those located along the
-the wind corridor. Three distinct variations has been detected.
+the wind corridor. Air particulate matter concentrations in urban-town
+sites of UB and DZ is episodically dictated by dust events in spring or
+late autumn, yet seasonally governed by anthropogenic emissions in
+winter. Air particulate matter concentrations in rural sites of SS and
+ZU is episodically dictated by dust events in spring or late autumn.
 
-Air quality in urban sites is episodically dictated by dust events in
-spring or late autumn, yet seasonally governed by anthropogenic
-emissions in winter. Air quality in rural sites of SS and ZU is
-episodically dictated by dust events in spring or late autumn. Air
-quality in rural sites of SS and ZU is episodically dictated by dust
-events in spring or late autumn.
+Three distinct variations has been detected.
 
-A clear seasonal variations in the sites of UB and DZ is {[}Air quality
-is governed by natural dust emission, and anthropogenic emissions{]}
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  It is evident of the new emission patterns in Mongolia.
+\item
+  Related to the winter emission patterns, fine particulates fraction in
+  the spring is increased.
+\item
+  This alarms\ldots{}
+\item
+  National level; meteorological impact is large. \ldots{} However,
+  reduced\ldots{} On the other hand, it is \ldots{} with the towns. This
+  points that air quality will be poor whether it is changed fuel, ..
+  unless to change heating system. Therefore, it is not the reason to
+  move the capital city. Only solution is to change the heating system,
+  do not burn any type of the coal.
+\end{enumerate}
 

### 2024-11-25 submission/Outline_20241121.Rmd
outline -  abstract, justify

@@
 Driven by mining industrial progress, household coal combustion has increased greatly in Mongolia since the 2000s.
-Demographic evidence has revealed an ongoing reduction in rural-nomadic lifestyle and a rise of the population at urban areas, which may gradually extend a household major source of fine particulate matter (PM2.5) air pollution in winter as a heating.
-Together with observations that lifespan in various animal species is flexible and can be increased by genetic or pharmaceutical intervention, these results have led to suggestions that longevity may not be subject to strict, species-specific genetic constraints.
+Demographic evidence has revealed an ongoing reduction in rural-nomadic lifestyle and a rise of the population at urban areas, which may gradually extend a household winter heating as a major source of fine particulate matter (PM2.5).
+Together ff with observations that lifespan in various animal species is flexible and can be increased by genetic or pharmaceutical intervention, these results have led to suggestions that longevity may not be subject to strict, species-specific genetic constraints.
 Here, by analysing global demographic data, we show that improvements in survival with age tend to decline after age 100, and that the age at death of the world’s oldest person has not increased since the 1990s.

### 2024-11-25 submission/Outline_20241121.tex
outline -  abstract, justify

@@
 the population at urban areas, which may gradually extend a household
-major source of fine particulate matter (PM2.5) air pollution in winter
-as a heating. Together with observations that lifespan in various animal
-species is flexible and can be increased by genetic or pharmaceutical
-intervention, these results have led to suggestions that longevity may
-not be subject to strict, species-specific genetic constraints. Here, by
-analysing global demographic data, we show that improvements in survival
-with age tend to decline after age 100, and that the age at death of the
-world's oldest person has not increased since the 1990s. Our results
-strongly suggest that the maximum lifespan of humans is fixed and
-subject to natural constraints. Storyline:
+winter heating as a major source of fine particulate matter (PM2.5).
+Together ff with observations that lifespan in various animal species is
+flexible and can be increased by genetic or pharmaceutical intervention,
+these results have led to suggestions that longevity may not be subject
+to strict, species-specific genetic constraints. Here, by analysing
+global demographic data, we show that improvements in survival with age
+tend to decline after age 100, and that the age at death of the world's
+oldest person has not increased since the 1990s. Our results strongly
+suggest that the maximum lifespan of humans is fixed and subject to
+natural constraints. Storyline:
 
